---
title: "Motif_network"
author: "Anneke Kakebeen"
date: "4/5/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---


## Set up
```{r setup, cache=FALSE, warning=FALSE, message=FALSE, echo=TRUE, error=TRUE}

# load packages
library(ggplot2)
library(LaCroixColoR)
library(RColorBrewer)
library(cowplot)
library(DT)
library(dplyr)
library(gridExtra)
library(tidyverse)
library(ggplotify)
library(knitr)
library(treemap)
library(gprofiler2)

# source functions
source("~/Desktop/pax6 paper/Final Markdowns/common_source_functions.R")

# Set wd
setwd("~/Desktop/pax6 paper/Final Markdowns/GRN/")

```


## Load Data
### DA ATAC
```{r DA ATAC, cache=TRUE, warning=FALSE, message=FALSE, echo=TRUE, error=TRUE}

# read in DA timepoints ATAC table
atac <- read.table(file="DA_timepoints.txt",
            sep="\t", header = TRUE, row.names = 1)
# shorten ATAC table to only pax
atac <- atac[,c(21,33,36,37, 38, 41,82:87)]
nrow(atac)

```

### RNASeq data
```{r Chang et al RNA-Seq, warning=FALSE}

# Read in Chang et al rna-seq
rna <- read.csv("RNA-Seq_DE_fuzzyk8_masterlist.csv")
nrow(rna)


# wrangle rna-seq
rna <- rna[,1:12] # remove rlog
rna <- rna[,-1*c(12,9)] # remove 15hpa
rna$sum <- rowSums(rna[,6:10]) # sum average log
rna <- filter(rna, rna$sum > 10.5) # use the sum as a filter to get out lowly expressed genes
remove <- grep(rna[,2], pattern = "LOC|Xetrov", ignore.case = TRUE) # set up vector to remove unnanotate
rna <- rna[-1*remove,] # remove unannotated genes
nrow(rna)

```

### motifs
```{r homer motifs, warning=FALSE}

# Write function to read in multiple tables
read.multi <- function(paths, outfiles){
  for (i in 1:length(paths)) {
   temp <- read.table(paths[i], sep = "\t", header = TRUE)
   assign(outfiles[i], temp, envir = .GlobalEnv)
  }
}

# Make inputs for function
motif_files <- list.files("annotate_outs/")
motif_paths <- paste("~/Desktop/pax6 paper/Final Markdowns/GRN/annotate_outs/", motif_files, sep = "")
motif_outs <- gsub(motif_files, pattern = ".bed.txt", replacement = "")

# Run funtion and make large motif table
read.multi(motif_paths, motif_outs)
object <- lapply(motif_outs, get)
# make motifs table
motifs <- do.call(what = dplyr::bind_rows, args = object)
nrow(motifs)

# wrangle motifs
# Extract TF Name into new column 
motifs$TF <- gsub("/.*","", motifs$Motif.Name)
motifs$TF <- gsub("\\(.*","", motifs$TF)
nrow(motifs)

```

## Integrate ATAC-RNA-Motif
```{r integrate data, warning=FALSE}
# Merge peaks motif files with DA table
Final.table <- merge.data.frame(motifs, atac, by.x = "PositionID", by.y = "Peak")
# make TF names lowercase
Final.table.short <- Final.table
Final.table.short$TF <- tolower(Final.table$TF)
# get rid of duplicates
Final.table.short = Final.table.short %>% distinct()
# get rid of unannotated genes
remove <- grep(Final.table.short$Gene, pattern = "LOC.*|Xetrov.*")
Final.table.short <- Final.table.short[-remove,]
# sort tf alphabetically
Final.table.short <- Final.table.short[order(Final.table.short$TF),]
nrow(Final.table.short)

# TFs that are expressed via RNASeq
tf.rna <- intersect(rna$Gene, Final.table.short$TF) # idenify TFs that are expressed in rna
tf.rna <- concat.grep(tf.rna) # convert to grep form
Final.table.short <- Final.table.short[grep(Final.table.short$TF, pattern = tf.rna),] # reduce master table to tfs expressed
nrow(Final.table.short)
```

## Shorten tables to only peaks that are DA at 6hpa, 24hpa, and 72hpa
### Table for 6hpa
```{r 6hpa, warning=FALSE}
# Make 6hpa table
table.6.1 <- rownames(Final.table.short[Final.table.short$DE.Pax_24hpaVSPax_6hpa == "DOWN",]) # 0< 6
table.6.2 <- rownames(Final.table.short[Final.table.short$DE.Pax_0hpaVSPax_6hpa == "DOWN",]) # 6> 24
table.6 <- unique(c(table.6.1, table.6.2)) # unique rows
table.6 <- Final.table.short[table.6,] 
table.6 <- na.omit(table.6) # remove NA
table.6$timepoint <- "6hpa" # add timepoint identifier for later
nrow(table.6)
```

### Table for 24hpa
```{r 24hpa, warning=FALSE}
table.24.1 <- rownames(Final.table.short[Final.table.short$DE.Pax_24hpaVSPax_6hpa == "UP",])
table.24.2 <- rownames(Final.table.short[Final.table.short$DE.Pax_24hpaVSPax_72hpa == "UP",])
table.24 <- unique(c(table.24.1, table.24.2))
table.24 <- Final.table.short[table.24,]
table.24 <- na.omit(table.24)
table.24$TF <- tolower(table.24$TF)
table.24$timepoint <- "24hpa"
nrow(table.24)
```

### Table for 72hpa
```{r 72hpa, warning=FALSE}
table.72.1 <- rownames(Final.table.short[Final.table.short$DE.Pax_72hpaVSPax_Whole.Tail == "UP",])
table.72.2 <- rownames(Final.table.short[Final.table.short$DE.Pax_24hpaVSPax_72hpa == "DOWN",])
table.72 <- unique(c(table.72.1, table.72.2))
table.72 <- Final.table.short[table.72,]
# sort for flags for 24hpa
table.72 <- na.omit(table.72)
table.72$TF <- tolower(table.72$TF)
table.72$timepoint <- "72hpa"
nrow(table.72)
```

### Table for uninjured v 24
```{r Uv24hpa, warning=FALSE}

table.24.1 <- rownames(Final.table.short[Final.table.short$DE.Pax_24hpaVSPax_Whole.Tail == "UP",])
table.u.24 <- Final.table.short[table.24.1,]

```

## Filter for motif scores
```{r filter for motif scores, warning=FALSE}

# save out prefiltered
write.table(table.6, "outs/table.6.txt", sep = "\t")
write.table(table.24, "outs/table.24.txt", sep = "\t")
write.table(table.72, "outs/table.72.txt", sep = "\t")

# motif scores were averaged over a fiven timepoint and all motifs below the average were excluded from the analysis
table.6.grn <- filter(table.6, MotifScore > mean(table.6$MotifScore))
table.24.grn <- filter(table.24, MotifScore > mean(table.24$MotifScore))
table.72.grn <- filter(table.72, MotifScore > mean(table.72$MotifScore))

write.table(table.6.grn, "outs/table.6.grn.txt", sep = "\t")
write.table(table.24.grn, "outs/table.24.grn.txt", sep = "\t")
write.table(table.72.grn, "outs/table.72.grn.txt", sep = "\t")
```

## Make one master table
```{r master table, warning=FALSE}

# Bind all DA tables to one 
mastertable <- rbind(table.24, table.6, table.72)
nrow(mastertable)

write.table(mastertable, "outs/atac_rna_motif.txt", sep = "\t")
mastertable <- read.table("outs/atac_rna_motif.txt", sep = "\t")


```


## Find timepoint gene/TF overlaps
### Read in tables if not continuing from above
```{r grn tables, warning=FALSE}

# table.6 <- read.table("outs/table.6.txt", sep = "\t")
# table.24 <- read.table("outs/table.24.txt", sep = "\t")
# table.72 <- read.table("outs/table.72.txt", sep = "\t")
# 
# 
# table.6.grn <- read.table("outs/table.6.grn.txt", sep = "\t")
# table.24.grn <- read.table("outs/table.24.grn.txt", sep = "\t")
# table.72.grn <- read.table("outs/table.72.grn.txt", sep = "\t")
# 
# mastertable <- rbind(table.24, table.6, table.72)


```

### predict GRN
```{r grn from filtered lists, warning=FALSE}

# looking to find GRN over 6->72hpa
# 6hpa DA region = 24hpa motif
# 24hpa DA region = 72hpa motif

# define 6 to 24hpa targets
grn.6.24 <- intersect(as.character(table.6.grn$Gene), table.24.grn$TF) # 6hpa DA region = 24hpa motif
grep.6.24 <- concat.grep(grn.6.24) # put into grp form

# define 24hpa to 72hpa targets
grn.24.72 <- intersect(table.24.grn$Gene, table.72.grn$TF) # 24hpa DA region = 72hpa motif
grep.24.72 <- concat.grep(grn.24.72)

# function to match 6->24 targets to 24-> 72hpa targets
grn.search <- function(table, genes, search, outs){
  for(i in 1:length(genes)){
    temp.table <- filter(table, TF == genes[i])
    intersect <- grep(temp.table[,17], pattern = search, value = TRUE, ignore.case = TRUE)
    assign(outs[i], intersect, envir = .GlobalEnv)
  }
}

# GRN across 6 to 72hpa
outs <- paste(grn.6.24, "out", sep = ".")
grn.search(table.24, grn.6.24,grep.24.72, outs) 

# GRN within 6hpa
outs <- paste(grn.6.24, "out.2", sep = ".")
grn.search(table.6, grn.6.24,grep.6.24, outs)

# GRN within 72hpa
outs <- paste(grn.24.72, "out.3", sep = ".")
grn.search(table.72, grn.24.72, grep.24.72, outs)

# GRN witin 24hpa
outs <- paste(grn.24.72, "out.4", sep = ".")
grn.search(table.24, grn.6.24, grep.6.24, outs)



```


## Identify other inputs and outs of pbx3 and meis 1
```{r pbx3 and meis1, warning=FALSE}

# filter master table for rows where TF is pbx3 and the gene is pbx3
Master_pbx <- filter(mastertable, TF == "pbx3" | Gene =="pbx3")
# filter master table for rows where TF is meis1 and the gene is meis1
Master_meis <- filter(mastertable, TF == "meis1" | Gene =="meis1")
# rowbind to one table
Master_pbx_meis <- rbind(Master_pbx, Master_meis)

# view table
datatable(Master_pbx_meis)


```

### pbx3 targets 1
```{r pbx3 targets 1, warning=FALSE}

# 6hpa
pbx3_6 <- as.character(filter(Master_pbx, timepoint=="6hpa" & TF == "pbx3")[,1])
length(pbx3_6)
# 24hpa
pbx3_24 <- as.character(filter(Master_pbx, timepoint=="24hpa" & TF == "pbx3")[,1])
length(pbx3_24)
# 72hpa
pbx3_72 <- as.character(filter(Master_pbx, timepoint=="72hpa" & TF == "pbx3")[,1])
length(pbx3_72)

pbx_all <- unique(c(pbx3_6, pbx3_24, pbx3_72))

#GO BP#
GO_pbx_terms <- gost(query = as.character(pbx_all), 
                    organism = "hsapiens",
                    sources = "GO:BP",
                    evcodes = TRUE,
                    significant = FALSE)  ##input gene list, run GO

# Make GO Table
GO_pbx_terms <- GO2table(GO_pbx_terms)

# Make heatmap for pbx3 targets
gene <- concat.grep(pbx3_72)
GOGenes <- grep(gene, timepoints$Peak, ignore.case = T) # get gene name rows from table

# Create heatmap matrix from counts data and input genes
paxGoGenes <- timepoints[GOGenes, ]
paxGoGenes <- paxGoGenes[-grep(paxGoGenes$Gene, pattern = "Xetrov|LOC", ignore.case = TRUE),] # remove unannotated
paxGoGenes.cpm <- log.counts.timepoints[rownames(paxGoGenes),]
rownames(paxGoGenes.cpm) <- paxGoGenes$Gene
paxGoGenes.cpm <- data.matrix(paxGoGenes.cpm)
paxGoGenes.cpm <- t(scale(t(paxGoGenes.cpm), scale = T))

colnames(paxGoGenes.cpm) <- gsub("_[0-9]$", "", colnames(paxGoGenes.cpm))
colnames(paxGoGenes.cpm) <- gsub("_2.1$", "", colnames(paxGoGenes.cpm))

mcounts <- t(apply(paxGoGenes.cpm, 1, function(x) {
  tapply(x, factor(colnames(paxGoGenes.cpm)), mean)
}))

colnames(mcounts) # view colnames to pick columns and order for heatmap
mcounts <- (mcounts[,c(6, 8, 7, 9 , 10)])

# Plot heatmap     
heatmap.2((mcounts), 
          scale="none",
          trace="none",
          Colv = FALSE,
          Rowv = TRUE, 
          dendrogram = 'none',
          density.info=c("none"),
          col=rev(color.palette),
          key = FALSE,
          cexRow=2, 
          cexCol=1, 
          labCol = FALSE, margins = c(0,10)) 

```

### pbx3  2
```{r pbx3 targets 2, warning=FALSE}

# 6hpa
pbx3_6 <- as.character(filter(Master_pbx, timepoint=="6hpa" & Gene == "pbx3")[,1])

# 24hpa
pbx3_24 <- as.character(filter(Master_pbx, timepoint=="24hpa" & Gene == "pbx3")[,1])

# 72hpa
pbx3_72 <- as.character(filter(Master_pbx, timepoint=="72hpa" & Gene == "pbx3")[,1])

pbx_all_genes <- unique(c(pbx3_6, pbx3_24, pbx3_72))

#GO BP#
GO_pbx_terms <- gost(query = as.character(pbx_all), 
                    organism = "hsapiens",
                    sources = "GO:BP",
                    evcodes = TRUE,
                    significant = FALSE)  ##input gene list, run GO

# Make GO Table
GO_pbx_terms <- GO2table(GO_pbx_terms)


```

### meis1 targets
```{r meis targets, warning=FALSE}

# 6hpa
meis1_6 <- as.character(filter(Master_meis, timepoint=="6hpa" & TF == "meis1")[,1])
length(meis1_6)

# 24hpa
meis1_24 <- as.character(filter(Master_meis, timepoint=="24hpa" & TF == "meis1")[,1])
meis1_24 <- filter(Master_meis, timepoint=="24hpa" & TF == "meis1")
meis1_24 <- filter(Master_meis, MotifScore > mean(Master_meis$MotifScore))
length(meis1_24)

# 72hpa
meis1_72 <- as.character(filter(Master_meis, timepoint=="72hpa" & TF == "meis1")[,1])
length(meis1_72)

meis_all <- unique(c(meis1_6, meis1_24, meis1_72))

#GO BP#
GO_meis_terms <- gost(query = as.character(meis_all), 
                    organism = "hsapiens",
                    sources = "GO:BP",
                    evcodes = TRUE)  ##input gene list, run GO

# Make GO Table
GO_meis_terms <- GO2table(GO_meis_terms)


# Make heatmap for pbx3 targets
gene <- concat.grep(meis1_24)
GOGenes <- grep(gene, timepoints$Peak, ignore.case = T) # get gene name rows from table

# Create heatmap matrix from counts data and input genes
paxGoGenes <- timepoints[GOGenes, ]
paxGoGenes <- paxGoGenes[-grep(paxGoGenes$Gene, pattern = "Xetrov|LOC|-like", ignore.case = TRUE),] # remove unannotated
paxGoGenes <- paxGoGenes[1:20,]
paxGoGenes.cpm <- log.counts.timepoints[rownames(paxGoGenes),]
rownames(paxGoGenes.cpm) <- paxGoGenes$Gene
paxGoGenes.cpm <- data.matrix(paxGoGenes.cpm)
paxGoGenes.cpm <- t(scale(t(paxGoGenes.cpm), scale = T))

colnames(paxGoGenes.cpm) <- gsub("_[0-9]$", "", colnames(paxGoGenes.cpm))
colnames(paxGoGenes.cpm) <- gsub("_2.1$", "", colnames(paxGoGenes.cpm))

mcounts <- t(apply(paxGoGenes.cpm, 1, function(x) {
  tapply(x, factor(colnames(paxGoGenes.cpm)), mean)
}))

colnames(mcounts) # view colnames to pick columns and order for heatmap
mcounts <- (mcounts[,c(6, 8, 7, 9 , 10)])

# Plot heatmap     
heatmap.2((mcounts), 
          scale="none",
          trace="none",
          Colv = FALSE,
          Rowv = TRUE, 
          dendrogram = 'none',
          density.info=c("none"),
          col=rev(color.palette),
          key = FALSE,
          cexRow=2, 
          cexCol=1, 
          labCol = FALSE, margins = c(0,10)) 

```



## motif treegraphs
```{r treegraph motifs, warning=FALSE}
# plot treemaps of TF inputs to genes and TF targets
library(treemap)

# meis -> etv1/runx ->
# inputs to meis1 expression at 6hpa
meis.6 <- filter(table.6, Gene == "meis1")
treegraph(meis.6, index = c("Gene", "TF"), directed = TRUE, show.labels = TRUE, vertex.label.cex = 1, vertex.label.color = "black", edge.arrow.mode=T)# edge.arrow.mode=T changes direction of arrows

# targets of meis1 TF at 24hpa
meis.24 <- filter(table.24, TF == "meis1")
treegraph(meis.24, index = c("TF", "Gene"), directed = TRUE, show.labels = TRUE, vertex.label.cex = 1, vertex.label.color = "black")

# runx1 in at 24hpa
runx1.in <- filter(table.24, Gene == "runx1")
treegraph(runx1.in, index = c("Gene", "TF"), directed = TRUE, show.labels = TRUE, vertex.label.cex = 1, vertex.label.color = "black", edge.arrow.mode=T)# edge.arrow.mode=T changes direction of arrows

# targets of runx1 at 72hpa
runx1 <- filter(table.72, TF == "runx1")
treegraph(runx1, index = c("TF", "Gene"), directed = TRUE, show.labels = TRUE, vertex.label.cex = 1, vertex.label.color = "black")

# runx2 in at 24hpa
runx2.in <- filter(table.24, Gene == "runx2")
treegraph(runx2.in, index = c("Gene", "TF"), directed = TRUE, show.labels = TRUE, vertex.label.cex = 1, vertex.label.color = "black", edge.arrow.mode=T)# edge.arrow.mode=T changes direction of arrows

# targets of runx2 at 72hpa
runx2 <- filter(table.72, TF == "runx2")
treegraph(runx2, index = c("TF", "Gene"), directed = TRUE, show.labels = TRUE, vertex.label.cex = 1, vertex.label.color = "black")

# etv1 in at 24hpa
etv1.in <- filter(table.24, Gene == "etv1")
treegraph(etv1.in, index = c("Gene", "TF"), directed = TRUE, show.labels = TRUE, vertex.label.cex = 1, vertex.label.color = "black", edge.arrow.mode=T)# edge.arrow.mode=T changes direction of arrows

# targets of etv1 at 72hpa
etv1 <- filter(table.72, TF == "etv1")
treegraph(etv1, index = c("TF", "Gene"), directed = TRUE, show.labels = TRUE, vertex.label.cex = 1, vertex.label.color = "black")



# etv1 ->pbx3 ->
# inputs to etv1 at 6hpa
etv1.6 <- filter(table.6, Gene == "etv1")
treegraph(etv1.6, index = c("Gene", "TF"), directed = TRUE, show.labels = TRUE, vertex.label.cex = 1, vertex.label.color = "black", edge.arrow.mode=T)# edge.arrow.mode=T changes direction of arrows

# targets of etv1 at 24hpa
etv1.24 <- filter(table.24, TF == "etv1")
treegraph(etv1.24, index = c("TF", "Gene"), directed = TRUE, show.labels = TRUE, vertex.label.cex = 1, vertex.label.color = "black")

# pbx3 in at 24hpa
pbx3.in <- filter(table.24, Gene == "pbx3")
treegraph(pbx3.in, index = c("Gene", "TF"), directed = TRUE, show.labels = TRUE, vertex.label.cex = 1, vertex.label.color = "black", edge.arrow.mode=T)# edge.arrow.mode=T changes direction of arrows

# targets of pbx3 at 72hpa
pbx3.out <- filter(table.72, TF == "pbx3")
treegraph(pbx3.out, index = c("TF", "Gene"), directed = TRUE, show.labels = TRUE, vertex.label.cex = 1, vertex.label.color = "black")



```


## Are meis1 or pbx3 differentially expressed?
```{r DE at all, warning=FALSE}

# read in single cell RNA-Seq DE data from uninj to 24hpa (upreg in 24hpa)
scDE <- read.table("~/Desktop/pax6 paper/Final Markdowns/SingleCell/analysis/outs/DEConditions_gain.txt", sep = "\t", header = TRUE, row.names  = 1)

# get rows for pbx3 and meis1
scDE_meis_pbx <- scDE[grep(scDE$gene, pattern = "meis1|pbx3"),]

# view rows
datatable(scDE_meis_pbx)

# How are pbx3 and meis1 expressed in bulk data
rna_meis_pbx <- rna[grep(rna$Gene, pattern = "meis1|pbx3"),] # filter table for pbx3 and meis1 rows
datatable(rna_meis_pbx) # view table

# make rna plot table
plot.rna <- as.data.frame(t(rna_meis_pbx[,6:10])) # transpose rna dataframe
colnames(plot.rna) <- c("measure", "measure")
plot.rna$timepoint <- gsub(rownames(plot.rna), pattern = "Mean_rlog_", replacement = "") # make column for timepoints
plot.rna <- rbind(as.data.frame(plot.rna[,c(1,3)]), as.data.frame(plot.rna[,2:3])) # make one measurement column
plot.rna$gene <- c(rep("pbx3", 5), rep("meis1",5)) # make gene column 
plot.rna$timepoint <- factor(plot.rna$timepoint, levels = c("WT", "0hpa", "6hpa", "24hpa", "72hpa"))

# plot data
ggplot(plot.rna, aes(x=timepoint, y=measure, color=gene, group=gene)) + # read in data
  geom_point() + # make scatter plot
  geom_line() + # connect points with lines
  facet_grid("gene") # facet by gene


```

## What GO terms do meis1 and pbx3 call
### Read in GO
```{r go terms, warning=FALSE}

# Read in GO Tables
Go6hpa <-  read.table("~/Desktop/pax6 paper/Final Markdowns/ATAC/ATAC_DA/outs_figures/GO6hpa.txt", sep = "\t", header = TRUE, row.names = 1)
Go6hpa$Timepoint <- "6hpa"
Go24hpa <-  read.table("~/Desktop/pax6 paper/Final Markdowns/ATAC/ATAC_DA/outs_figures/GO24hpa.txt", sep = "\t", header = TRUE, row.names = 1)
Go24hpa$Timepoint <- "24hpa"
Go72hpa <-  read.table("~/Desktop/pax6 paper/Final Markdowns/ATAC/ATAC_DA/outs_figures/GO72hpa.txt", sep = "\t", header = TRUE, row.names = 1)
Go72hpa$Timepoint <- "72hpa"

# rbind tables
Go <- rbind(Go6hpa, Go24hpa, Go72hpa)
```

### meis1
```{r meis1 GO, warning=FALSE}
# look for meis
Go_meis <- Go[grep(Go$intersection, pattern = "meis1"),]

# view table
datatable(Go_meis)

```

### pbx3
```{r pbx3 GO}
# look for meis
Go_pbx3 <- Go[grep(Go$intersection, pattern = "pbx3"),]

# view table
datatable(Go_pbx3)

```




