---
title: "PH3 Cells in Regen"
author: "Anneke Kakebeen"
date: "8/14/2019 (Revised 09/27/19)"
always_allow_html: yes
output:
  github_document:
    toc: true
    toc_depth: 3
---

## Set up 
```{r set up, echo=FALSE, warning=FALSE, message=FALSE}

# load packages 
library(ggplot2)
library(tidyr)
library(DT)
library(LaCroixColoR)
library(dplyr)
library(ggpubr)
library(cowplot)
library(Seurat)
library(patchwork)

# colors 
meisColor <- "#FBB58F"
pbxColor <- "#B574A7"
color <- lacroix_palette("CranRaspberry")

# source functions
source("~/Desktop/R_working/common_source_functions.R")


# set wd
setwd("~/Desktop/pax6 paper/Final Markdowns/RegenPH3/")

```

## Load in data
```{r load data, echo=FALSE, warning=FALSE, cache=TRUE}

# read in data
ph3 <- read.csv("regen_ph3.csv")


# view data
# datatable(ph3)

```

## Figure 5AA: PH3 cells in 24hpa and 72hpa tissue
### subset working data
```{r Figure5AA data, echo=FALSE, warning=FALSE, cache=TRUE}

# subset data for plot
ph3_2472 <- filter(ph3, Clutch == "20720")

# save source file
# write.csv(ph3_2472, "ph3_24hpa_72hpa.csv")
```

### Plot 24hpa and 72hpa
Conclusion: 72hpa has significant more ph3 cells per area of spinal cord.
```{r Figure5AA plot, echo=FALSE, warning=FALSE, cache=TRUE}
# plot ph3/area of spinal cord FOR FIGURES
ggplot(ph3_2472, aes(x=Timepoint, y=CPMm)) +
  geom_boxplot(fill="green4") +
  ylab(expression("pH3 Cells per"~mm^2~" of the Spinal Cord")) +
  theme_bw() +
  theme(text = element_text(colour = "black", size = 29), axis.text.x = element_text(color="black"), axis.text.y = element_text(color="black")) +
  stat_compare_means(comparisons = list(c("24hpa", "72hpa")), method = "t.test") 
 

# plot ph3/area of spinal cord FOR TALKS
# ggplot(ph3, aes(x=Timepoint, y=CPMm)) +
#   geom_boxplot(fill="green4", color="white") +
#   DarkTheme() +
#   ylab(expression("Cells per"~mm^2~" of the Spinal Cord")) +
#   theme(text = element_text(colour = "white")) +
#   stat_compare_means(comparisons = list(c("24hpa", "72hpa")), method = "t.test", label = "p.signif", color = "white") 

# number of animals for each condition
table(ph3_2472$Timepoint)

```


## Figure 7L-M: Area of regenerated spinal cord and ph3 per area
### Subset working data
```{r Figure7L-M data, echo=FALSE, warning=FALSE, cache=TRUE}
#filter for morphant data
ph3_morphant <- filter(ph3, Clutch == "22120")

# source for morphant data
# write.csv(ph3_morphant, "ph3_morphant.csv")

# Factor conditions for plotting
ph3_morphant$Condition <- factor(ph3_morphant$Condition, levels = c("tracer", "meis1_VMO", "pbx3_VMO"))
```

### plot morphant data
Conclusion: There is a significant decrease in area of regen spinal cord in morphants, howver the number of ph3 cells per area does not change in morphant conditions.
```{r Figure7L-M daplotta, echo=FALSE, warning=FALSE, cache=TRUE}

# set stats covmparison
comparison <- list(c("tracer", "pbx3_VMO"), c("tracer", "meis1_VMO"))

# Plot area of regen spinal cord
ggplot(subset(ph3_morphant, Timepoint == "72hpa" & Condition %in% c("tracer", "meis1_VMO", "pbx3_VMO")), aes(x=Condition, y=(Area_micron/1000))) +
  geom_boxplot(fill="green4") +
  ylab(expression("Area of Spinal Cord"~paste(mm^2))) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text = element_text(colour = "black", size = 29), axis.text.x = element_text(color="black"), axis.text.y = element_text(color="black")) +
  stat_compare_means(comparisons = comparison, method = "t.test", label = "p.signif") +
  xlab("") +
  Seurat:::RotatedAxis()

# plot ph3 cells per area spinal cord
ggplot(subset(ph3_morphant, Timepoint == "72hpa"& Condition %in% c("tracer", "meis1_VMO", "pbx3_VMO")), aes(x=Condition, y=CPMm)) +
  geom_boxplot(fill="green4") +
  ylab(expression("PH3 Cells/"~mm^2~"Spinal Cord")) +
  theme_bw() +
  xlab("") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text = element_text(colour = "black", size = 29), axis.text.x = element_text(color="black"), axis.text.y = element_text(color="black")) +
  stat_compare_means(comparisons = comparison, method = "t.test", label = "p.signif") +
  Seurat:::RotatedAxis()


table(ph3_morphant$Condition, ph3_morphant$Timepoint)


```