---
title: "Tail Measurements"
author: "Anneke Kakebeen"
date: "8/14/2019 (Revised 09/27/19)"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Set up 
```{r set up, echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE}

# load packages 
library(ggplot2)
library(tidyr)
library(DT)
library(LaCroixColoR)
library(dplyr)
library(ggpubr)
library(cowplot)
library(Seurat)

# colors 
uninjColor <- "white"
tracerColor <- "#D9565C"
meisColor <- "#FBB58F"
pbxColor <- "#B574A7"
color <- lacroix_palette("CranRaspberry")

# source functions
source("~/Desktop/R_working/common_source_functions.R")


# set wd
setwd("~/Desktop/pax6 paper/Final Markdowns/TailMeasurement/")

```

## Load in data
This .csv file has values for all regeneration measurements taken in this study. Plots will be made on derivatives of this data. The working file for each chunck will be saved to submit to Elife as source file for data. 
```{r load data, echo=FALSE, warning=FALSE, cache=TRUE}

# read in data
MOdata <- read.csv("pbx3_meis1MO_regenmeasures_021220.csv")
MOdata$SampleFull <- paste(MOdata$Sample, MOdata$Timepoint, MOdata$Replicate, MOdata$Clutch, sep = ".") # make column with sampleID
MOdata$Clutch <- as.character(MOdata$Clutch) # set numerical values of clutch as a character 

# phenotype definition
# morphology: 1=WT, 2= full epidermal covering, 3=neither, 4=axial nub, 5=axial no fin
# neurons: 10=neurons in regen, 20=no neurons in regen, 30=not stained
# notocord: 100=stacked, 200=not stacked

# rename sc to spinal cord
MOdata$MeasureType <- (gsub(MOdata$MeasureType, pattern = "sc", replacement = "Spinal Cord"))
MOdata$MeasureType <- factor(MOdata$MeasureType, levels = c("Tail", "Spinal Cord", "Spinal Cord_width")) # factor measurement type for plot 

# view data
datatable(MOdata)

```


## Figure 7K: Plot tail and spinal cord measurement data for vivo-morphants
### Subset working data for vmo
```{r Figure 7k data,  echo=FALSE, warning=FALSE, cache=TRUE}

# filter data for last clutch with correct doses
vmo_022120 <- MOdata %>%
  filter(Clutch == "22120" ) %>%
  filter(SampleID %in% c("tracer", "pbx3_VMO", "meis1_VMO"))

# write source table
# write.csv(vmo_022120, "vivo_morphant_tail_measurements.csv")

#factor data to order sample variables correctly
vmo_022120$SampleID <- factor(vmo_022120$SampleID, levels = c("tracer", "meis1_VMO", "pbx3_VMO"))
```

### Plot 72hpa for fig7k
Conclusions: meis1 and pbx3 vivo-morpholinos result in significantly shorter regenerated spinal cord and tail. 
```{r Figure 7k plot,  echo=FALSE, warning=FALSE, cache=TRUE}

# define comparisons for statistics
comparisons <- list(c("tracer", "meis1_VMO"), c("tracer", "pbx3_VMO")) # define comparisons for statistics

# for publication
ggplot(vmo_022120, aes(x=SampleID, y=Measurement, fill=SampleID)) +
  geom_boxplot() +
  theme_bw(base_size = 29) +
  theme(legend.position = "none", panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.text.x = element_text(colour = "black"), axis.text.y = element_text(colour = "black"), strip.background =element_rect(fill="white")) +
  facet_grid(Timepoint~MeasureType) +
  ylab(expression(paste("Length ", mu, "m"))) +
  xlab("") +
   stat_compare_means(comparisons = comparisons, method = "t.test", label = "p.signif") +
    Seurat:::RotatedAxis() +
  scale_fill_manual(values = c(tracerColor, meisColor, pbxColor))


# number of animals for each condition
table(vmo_022120$SampleID, vmo_022120$MeasureType)

```


## Figure 7-supp1AA: Plot tail measurement data for embryonically injected morphants
### Subset working data for embryonic
```{r fig7-supp1AA data, echo=FALSE, warning=FALSE, cache=TRUE}
 
# Get data for embryonic morphants
# filter table for embryonic samples and only tail and spinal cord measurements
TailS <- MOdata %>%
  filter(Other.Notes == "embryonic study") %>%
  filter(MeasureType %in% c("Tail", "Spinal Cord")) 

# write out working data for source file
# write.csv(TailS, "embryonic_morphant_tail_measurements.csv")
```

### Plot 72hpa for fig7-supp1AA 
Conclusions: meis1 and pbx3 morpholinos result in significantly shorter regenerated spinal cord and tail. 
```{r fig7-supp1AA plot, echo=FALSE, warning=FALSE, cache=TRUE}
# define comparisons for statistics
comparisons <- list(c("cntrl", "meis1MO1"),c("cntrl", "meis1MO2"),c("cntrl", "pbx3MO1"),c("cntrl", "pbx3MO2")) 

# plot 72hpa
ggplot(filter(TailS, Timepoint == "72hpa"), aes(x=SampleID, y=Measurement, fill=SampleID)) +
  geom_boxplot() +
  geom_jitter() +
  theme_bw(base_size = 20) +
  theme(legend.position = "none", panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.text.x = element_text(colour = "black"), axis.text.y = element_text(colour = "black")) +
  facet_grid(Timepoint~MeasureType) +
  ylim(0, 1000) +
  ylab(expression(paste("Length ", mu, "m"))) +
  xlab("") +
   stat_compare_means(comparisons = comparisons, method = "t.test", label.y = c(750, 825, 900, 975), label = "p.signif") +
    Seurat:::RotatedAxis() +
  scale_fill_manual(values = c("white",meisColor, meisColor, pbxColor, pbxColor))

# number of animals for each condition
TailS_72 <- filter(TailS, Timepoint == "72hpa")
table(TailS_72$SampleID, TailS_72$MeasureType)

```

### Plot all timepoints for reference
```{r all timepoints plot, echo=FALSE, warning=FALSE, cache=TRUE}


# plot all timepoints
ggplot(TailS, aes(x=SampleID, y=Measurement, fill=SampleID)) + # define dataframe and paramters
  geom_boxplot() + # make boxplot
  geom_jitter() + # add points on top of boxplots
  theme_bw(base_size = 20) + # adjust text size
  theme(legend.position = "none", panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + # adjust graph look (remove gridlines)
  facet_grid(Timepoint~MeasureType) + # divide graph by timepoint and measurement type
  ylim(0, 1000) + # define axis limit
  ylab(expression(paste("Length ", mu, "m"))) + # label y axis
  xlab("") + # remove xaxis title
   stat_compare_means(comparisons = comparisons, method = "t.test", label.y = c(750, 825, 900, 970), label = "p.signif") + # add statistical test
    Seurat:::RotatedAxis() + # rotate xaxis names
  scale_fill_manual(values = c("white",meisColor, meisColor, pbxColor, pbxColor)) # call fill colors


```


