---
title: "ATAC Figures"
author: "Anneke Kakebeen"
date: "13 Aug 2019"
output:
  html_document:
    code_folding: "hide"
    toc: true
    toc_depth: 3
    toc_float: true
fig_width: 6
fig_height: 4
editor_options: 
  chunk_output_type: console
---
## Setting up project: ATAC
```{r set up, cache=FALSE, warning=FALSE, message=FALSE}

# Load packages
library(knitr)
library(gridExtra)
library(dplyr)
library(gprofiler2)
library(ggplot2)
library(AnnotationDbi)
library(biomaRt)
library(gplots)
library(tidyverse)
library(RColorBrewer)
library(DT)
library(LaCroixColoR)
library(Seurat)
library(VennDiagram)
library(cowplot)
library(imager)
library(magick)
library(pdftools)
library(limma)
library(treemap)


# source functions
source("~/Desktop/R_working/common_source_functions.R")

## color palette for heatmaps
color.palette <- colorRampPalette(lacroix_palette("Lemon", n = 50, type = "continuous"))(256) 

# setwd
setwd("~/Desktop/pax6 paper/Final Markdowns/ATAC/ATAC_DA/")

```

## Read in data for figures
```{r read in data, cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE}

# read in DA tables
paxVall_agg <- read.table("outs/DA_paxVall_ALL.txt", header = TRUE, sep = "\t", row.names = 1) # paxVall aggregate
paxVall <- read.table("outs/DA_paxVall.txt", header = TRUE, sep = "\t", row.names = 1) # paxVall aggregate
timepoints <- read.table("outs/DA_timepoints.txt", header = TRUE, sep = "\t", row.names = 1) # between timepoints within one condition

# Read in log counts
log.counts.paxVall <- as.matrix(read.table("outs/log.counts_paxVall_aggregate.txt", header = TRUE, sep = "\t", row.names = 1)) # log counts for paxVall aggregate
log.counts.timepoints <- as.matrix(read.table("outs/log.counts_timepoints.txt", header = TRUE, sep = "\t", row.names = 1)) # log counts for timepoints

```



## Figure 2B
```{r 2b, cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE, fig.width= 8, fig.height= 8}

# read in mds data
targets <- read.table("~/Desktop/pax6 paper/Final Markdowns/ATAC/ATAC_DA/outs/targets.txt", sep = "\t")
mds <- readRDS("~/Desktop/pax6 paper/Final Markdowns/ATAC/ATAC_DA/outs/mds.RDS")

# assign color groups to pax and all-tissue
Group <- factor(paste(targets$CellID, targets$TimePt, sep="_"))
col.group <- as.character(Group) #grouped by tissue type, ie Pax or Whole.Tail
col.group <- gsub("All_Tissue", "All.Tissue", col.group) # adjust name for color usage
col.group <- gsub("_.*", "", col.group)
col.group <- gsub("All.Tissue", "royalblue3", col.group) #change Whole tail samples to blue
col.group <- gsub("Pax", "green4", col.group) # change pax samples to green

# plot MDS
plotMDS(mds,
        col = col.group,
        main = "MDS Plot")

```


## Figure 2C: tss plots
```{r 2c, cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE, fig.width= 8, fig.height= 8}
# Fix when alex gets back to you
# pax <- image_read("~/Desktop/Final_R/ATAC/tss_pax.png")
# print(pax)
# 
# all <- image_read("~/Desktop/Final_R/ATAC/tss_all.png")
# print(all)

```


## Figure 2E: # of peaks called DA between conditions
### Number of peaks
```{r peaks between conditions, cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE}

# Make datatable for plottig
# Get # of peaks called diffferential in each direction in contrasts of interest
Freq <- c(nrow(paxVall_agg[paxVall_agg$DE.p.PaxVSWholeALL == "UP",]), # pax > all
          nrow(paxVall_agg[paxVall_agg$DE.p.PaxVSWholeALL == "DOWN",]), # pax < all
          nrow(filter(paxVall, paxVall$DE.PaxVSWhole_WT == "UP")), # uninjured pax > all
          nrow(filter(paxVall, paxVall$DE.PaxVSWhole_WT == "DOWN")), # uninjured pax < all
          nrow(filter(paxVall, paxVall$DE.PaxVSWhole_0hpa == "UP")), # 0hpa pax > all
          nrow(filter(paxVall, paxVall$DE.PaxVSWhole_0hpa == "DOWN")), # 0hpa pax < all
          nrow(filter(paxVall, paxVall$DE.PaxVSWhole_6hpa == "UP")), # 6hpa pax > all
          nrow(filter(paxVall, paxVall$DE.PaxVSWhole_6hpa == "DOWN")), # 6hpa pax < all
          nrow(filter(paxVall, paxVall$DE.PaxVSWhole_24hpa == "UP")), # 24hpa pax > all
          nrow(filter(paxVall, paxVall$DE.PaxVSWhole_24hpa == "DOWN")), # 24hpa pax < all
          nrow(filter(paxVall, paxVall$DE.PaxVSWhole_72hpa == "UP")), # 72hpa pax > all
          nrow(filter(paxVall, paxVall$DE.PaxVSWhole_72hpa == "DOWN")), # 72hpa pax < all
          length(unique(paxVall_agg[paxVall_agg$DE.p.PaxVSWholeALL == "UP",12])), # pax > all
          length(unique(paxVall_agg[paxVall_agg$DE.p.PaxVSWholeALL == "DOWN",12])), # pax < all
          length(unique((paxVall[paxVall$DE.PaxVSWhole_WT == "UP", 26]))), # uninjured pax > all
          length(unique((paxVall[paxVall$DE.PaxVSWhole_WT == "DOWN",26]))), # uninjured pax < all
          length(unique((paxVall[paxVall$DE.PaxVSWhole_0hpa == "UP",26]))), # 0hpa pax > all
          length(unique((paxVall[paxVall$DE.PaxVSWhole_0hpa == "DOWN", 26]))), # 0hpa pax < all
          length(unique((paxVall[paxVall$DE.PaxVSWhole_6hpa == "UP",26]))), # 6hpa pax > all
          length(unique((paxVall[paxVall$DE.PaxVSWhole_6hpa == "DOWN",26]))), # 6hpa pax < all
          length(unique((paxVall[paxVall$DE.PaxVSWhole_24hpa == "UP",26]))), # 24hpa pax > all
          length(unique((paxVall[paxVall$DE.PaxVSWhole_24hpa == "DOWN",26]))), # 24hpa pax < all
          length(unique((paxVall[paxVall$DE.PaxVSWhole_72hpa == "UP",26]))), # 72hpa pax > all
          length(unique((paxVall[paxVall$DE.PaxVSWhole_72hpa == "DOWN",26]))) # 72hpa pax < all
          
          )

# Make metadata columns
Contrast <- c(rep(c("All Timepoints", "All Timepoints", "Uninjured", "Uninjured", "0hpa", "0hpa", "6hpa", "6hpa", "24hpa", "24hpa", "72hpa", "72hpa"),2)) # column for timepoint designation
Direction <- c(rep(rep(c("pax6", "all-tissue"), 6),2)) # column for favored in accessibltiy comparison
Measure <- c(rep("peaks",12), rep("genes",12))
peakbreak <- data.frame(Contrast, Freq, Direction, Measure) # make dataframe
peakbreak$Freq <- ifelse(peakbreak$Direction == "pax6", -1*peakbreak$Freq, peakbreak$Freq) # make pax6 frequency values negative for 

# order table for plotting by facotring
peakbreak$Contrast <- factor(peakbreak$Contrast, levels = (c("All Timepoints", "Uninjured", "0hpa", "6hpa", "24hpa", "72hpa"))) 
peakbreak$Direction <- factor(peakbreak$Direction, levels = c("pax6", "all-tissue"))

# Make frequency into percentage
# get totals per contrast and measurement type
total <- as.data.frame(peakbreak %>%
  group_by(Contrast, Measure) %>%
  summarize(total = sum(abs(Freq))))
# merge total and peak dataframes
peakbreak <- merge.data.frame(peakbreak, total, by = c("Contrast", "Measure"))
# make percentage column
peakbreak$Percent <- abs(peakbreak$Freq)/peakbreak$total*100


# view table
datatable(peakbreak, caption = "Number of peaks called in pax or all tissue libraries per timepoint")


# Plot pyramid chart of peaks 
ggplot(filter(peakbreak, Measure == "peaks"), aes(x= Contrast, y= Percent, fill=Direction))+
  geom_bar(stat="identity", width = 0.8)  +
  scale_fill_manual(values = c("forestgreen","dodgerblue3", "#F7AA14", "#172869")) +
  xlab("") +
  ylab("% of DA Peaks (all-tissue vs. pax6)") +
  theme_bw(base_size = 30) +
    Seurat:::RotatedAxis() +
    theme(legend.position = "none", axis.text = element_text(colour = "black")) 



# Plot pyramid chart of peaks and genes
ggplot((peakbreak), aes(x= Contrast, y= Freq, fill=Measure))+
  geom_bar(stat="identity", position = "dodge")  +
  # facet_wrap(~Contrast, ncol = 1, scales = "free_y") +
  # facet_grid(Contrast~., space = "free_x", scales = "free_y", switch = "x") +
  scale_fill_manual(values = c("forestgreen","dodgerblue3", "#F7AA14", "#172869")) +
  coord_flip() +
  ylim(-5500, 5500) +
  Seurat:::RotatedAxis() +
  theme(legend.position = "bottom") +
theme(strip.placement = "outside",
strip.background = element_rect(fill=NA,colour="white"),
panel.spacing.x=unit(0,"cm"),
strip.text = element_text(hjust=0, face="bold", size=12)) +
  theme_bw() +
  facet_grid(Measure~.)

```


## Figure 2F: GO analysis of aggregate pax v all regions
### PAX > All
```{r 2f.1 , cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE}

# Define input regions
GO_pax <- filter(paxVall_agg, paxVall_agg$DE.p.PaxVSWholeALL == "UP") # filter table for all regions with accessiblity pax > all
GO_pax <- GO_pax[,c(12, 3)] # get genes and flags

# view inputs for GO
datatable(GO_pax, caption = "regions more accessible in pax than all-tissue")

# Run gprofileR2
GO_out <- gost(query = as.character(GO_pax$Gene), 
                    organism = "hsapiens",
                    sources = "GO:BP",
                    evcodes = TRUE)  ##input gene list, run GO

# Make table of GO terms (Funciton in common source functions)
GO_PAX_use <- GO2table(GOResults = GO_out)

# View GO outputs
datatable(GO_PAX_use)

# Plot go terms in barplot
ggplot(GO_PAX_use[1:5,], aes(x= term_name, y= logp))  +
  geom_bar(stat = "identity", fill = "green4") +
  coord_flip()+
  labs(x = "GO Terms", y = "-log(pvalue)", title = "Pax6 Top GO Terms") +
  blank_theme

# write out table
write.table(GO_PAX_use, "~/Desktop/pax6 paper/Final Markdowns/ATAC/ATAC_DA/outs_figures/GO_PAX_use.txt", sep = "\t")

```

### ALL v pax GO
Goal : Show that regions that are more accessible in pax libraries vs all-tissue libraries have neural character
Method: Identify differential regions of the chromatin between Pax and AT, use these regions for GO
Conclusions: Regions prioritized in all tissue show more generic GO terms. 
```{r ALL , cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE}


# Define input regions
GO_all <- filter(paxVall_agg, paxVall_agg$DE.p.PaxVSWholeALL == "DOWN") # filter table for all regions with accessiblity pax < all
GO_all <- GO_all[,c(12, 3)] # get genes and flags

# view table
datatable(GO_all, caption = "regions more accessible in all than pax")


# Run gprofileR2
GO_out <- gost(query = as.character(GO_all$Gene), 
                    organism = "hsapiens",
                    sources = "GO:BP",
                    evcodes = TRUE)  ##input gene list, run GO

# Make table of GO terms (Funciton in common source functions)
GO_ALL_use <- GO2table(GOResults = GO_out)

# View GO outputs
datatable(GO_ALL_use)

# Plot go terms in barplot
ggplot(GO_ALL_use[1:5,], aes(x= term_name, y= logp))  +
  geom_bar(stat = "identity", fill = "dodgerblue3") +
  coord_flip()+
  labs(x = "GO Terms", y = "-log(pvalue)", title = "All Tissue Top GO Terms") +
  blank_theme

# write out table
write.table(GO_ALL_use, "~/Desktop/pax6 paper/Final Markdowns/ATAC/ATAC_DA/outs_figures/GO_ALL_use.txt", sep = "\t")

```

## Figure 2G: heat map of terms in neurogenesis
```{r PaxVAll, cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE}
# read in go terms
GO_PAX_use <- read.table("~/Desktop/pax6 paper/Final Markdowns/ATAC/ATAC_DA/outs_figures/GO_PAX_use.txt", header = TRUE, sep = " ")

# Define regions where npax > all
pax <- paxVall_agg[c(paxVall_agg$DE.p.PaxVSWholeALL == "UP", paxVall_agg$Gene == "pax6"),]

# define genes 
gene <- c(concat.GO(GO_PAX_use[7,4]), "pax6")
# gene <- c("SEMA3F,DVL2,PROM1,PAX6,ISL1,CNTN1,MUSK,USH2A,HDAC9,ARID1B") # list for "neurogenesis"
# print(gene)
gene <- concat.grep(gene)
GOGenes <- grep(gene, pax$Gene, ignore.case = T) # get gene name rows from table

# Create heatmap matrix from counts data and input genes
paxGoGenes <- pax[GOGenes, ] # get dataframe with genes from neurogenesis go term
paxGoGenes <- paxGoGenes[c(sample(nrow(paxGoGenes), 19), grep(paxGoGenes$Gene, pattern = "pax6")),] # randomly sample 20 rows for heatmap
paxGoGenes.cpm <- log.counts.paxVall[rownames(paxGoGenes),]
rownames(paxGoGenes.cpm) <- paxGoGenes$Gene
paxGoGenes.cpm <- data.matrix(paxGoGenes.cpm)
paxGoGenes.cpm <- t(scale(t(paxGoGenes.cpm), scale = T))

colnames(paxGoGenes.cpm) <- gsub("_[0-9]$", "", colnames(paxGoGenes.cpm))
colnames(paxGoGenes.cpm) <- gsub("_2.1$", "", colnames(paxGoGenes.cpm))

mcounts <- t(apply(paxGoGenes.cpm, 1, function(x) {
  tapply(x, factor(colnames(paxGoGenes.cpm)), mean)
}))

# colnames(mcounts) # view colnames to pick columns and order for heatmap
mcounts <- (mcounts[,c(6,8,7,9,10,1,3,2,4,5)])

# Plot heatmap     
heatmap.2(t(mcounts), 
          scale="none",
          trace="none",
          Colv = TRUE,
          Rowv = FALSE, 
          dendrogram = 'none',
          density.info=c("none"),
          col=rev(color.palette),
          key = FALSE,
          cexCol=2, 
          labRow = FALSE,
          margins = c(10,2), 
          srtCol = 45 )   # angle of labels

```



## Figure 3A: heatmaps for all DA regions at 6hpa, 24hpa, 72hpa
### 6hpa Heatmap 
```{r 6hpa heat, cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE}
# Get regions
tss_6hpa_1_Peaks <- rownames(timepoints[timepoints$DE.Pax_0hpaVSPax_6hpa == "DOWN",]) # regions 6 > 0
tss_6hpa_2_Peaks <- rownames(timepoints[timepoints$DE.Pax_24hpaVSPax_6hpa == "DOWN",]) # regions 6 > 24
tss_6hpa_Peaks <- unique(c(tss_6hpa_1_Peaks, tss_6hpa_2_Peaks)) # unique peaks
tss_6hpa <- timepoints[tss_6hpa_Peaks,] # make table with only these peaks

# Make 6hpa heatmap matrix
paxGoGenes <- tss_6hpa[1:nrow(tss_6hpa),]
paxGoGenes.cpm <- log.counts.timepoints[rownames(paxGoGenes),]
rownames(paxGoGenes.cpm) <- paxGoGenes$Gene
paxGoGenes.cpm <- data.matrix(paxGoGenes.cpm)
paxGoGenes.cpm <- t(scale(t(paxGoGenes.cpm), scale = T))

# library(dplyr)
colnames(paxGoGenes.cpm) <- gsub("_[0-9]$", "", colnames(paxGoGenes.cpm))
colnames(paxGoGenes.cpm) <- gsub("_2.1$", "", colnames(paxGoGenes.cpm))

mcounts <- t(apply(paxGoGenes.cpm, 1, function(x) {
  tapply(x, factor(colnames(paxGoGenes.cpm)), mean)
}))

mcounts <- mcounts[,c(6,8,7,9,10)]
      
# plot heatmap        
heatmap.2(t(mcounts), 
          scale="none",
          trace="none",
          Colv = TRUE,
          Rowv = FALSE,
          dendrogram = 'none',
          main = "6hpa TSS REGIONS",
          density.info=c("none"),
          col=rev(color.palette),
          cexRow=0.7, cexCol=0.75, 
          labCol = FALSE) ##3takes 10 minues
```

### 24hpa Heatmap 
```{r 24hpa heat, cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE}

# Get regions
tss_24hpa_1_Peaks <- rownames(timepoints[timepoints$DE.Pax_24hpaVSPax_6hpa == "UP",]) # 24 > 6
tss_24hpa_2_Peaks <- rownames(timepoints[timepoints$DE.Pax_24hpaVSPax_72hpa == "UP",]) # 24 > 72
tss_24hpa_Peaks <- unique(c(tss_24hpa_1_Peaks, tss_24hpa_2_Peaks))
tss_24hpa <- timepoints[tss_24hpa_Peaks,]

# make heatmap matrix
paxGoGenes <- tss_24hpa[1:nrow(tss_24hpa),]
paxGoGenes.cpm <- log.counts.timepoints[rownames(paxGoGenes),]
rownames(paxGoGenes.cpm) <- paxGoGenes$Gene
paxGoGenes.cpm <- data.matrix(paxGoGenes.cpm)
paxGoGenes.cpm <- t(scale(t(paxGoGenes.cpm), scale = T))

# library(dplyr)
colnames(paxGoGenes.cpm) <- gsub("_[0-9]$", "", colnames(paxGoGenes.cpm))
colnames(paxGoGenes.cpm) <- gsub("_2.1$", "", colnames(paxGoGenes.cpm))

mcounts <- t(apply(paxGoGenes.cpm, 1, function(x) {
  tapply(x, factor(colnames(paxGoGenes.cpm)), mean)
}))

mcounts <- mcounts[,c(6,8,7,9,10)]
      
# plot heatmap        
heatmap.2(t(mcounts), 
          scale="none",
          trace="none",
          Colv = TRUE,
          Rowv = FALSE,
          dendrogram = 'none',
          main = "24hpa TSS REGIONS",
          density.info=c("none"),
          col=rev(color.palette),
          cexRow=0.7, cexCol=0.75, 
          labCol = FALSE) ##3takes 10 minues
```

### 72hpa Heatmap 
```{r 72hpa heat, cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE}

# Get regions
tss_72hpa_1_Peaks <- rownames(timepoints[timepoints$DE.Pax_24hpaVSPax_72hpa == "DOWN",]) # 72 > 24
tss_72hpa_2_Peaks <- rownames(timepoints[timepoints$DE.Pax_72hpaVSPax_Whole.Tail == "UP",]) # 72 > uninj
tss_72hpa_Peaks <- unique(c(tss_72hpa_1_Peaks, tss_72hpa_2_Peaks))
tss_72hpa <- timepoints[tss_72hpa_Peaks,]

# make heatmap matrix
paxGoGenes <- tss_72hpa[1:nrow(tss_72hpa),]
paxGoGenes.cpm <- log.counts.timepoints[rownames(paxGoGenes),]
rownames(paxGoGenes.cpm) <- paxGoGenes$Gene
paxGoGenes.cpm <- data.matrix(paxGoGenes.cpm)
paxGoGenes.cpm <- t(scale(t(paxGoGenes.cpm), scale = T))

# library(dplyr)
colnames(paxGoGenes.cpm) <- gsub("_[0-9]$", "", colnames(paxGoGenes.cpm))
colnames(paxGoGenes.cpm) <- gsub("_2.1$", "", colnames(paxGoGenes.cpm))

mcounts <- t(apply(paxGoGenes.cpm, 1, function(x) {
  tapply(x, factor(colnames(paxGoGenes.cpm)), mean)
}))

mcounts <- mcounts[,c(6,8,7,9,10)]
      
# plot heatmap        
heatmap.2(t(mcounts), 
          scale="none",
          trace="none",
          Colv = TRUE,
          Rowv = FALSE,
          dendrogram = 'none',
          main = "72hpa TSS REGIONS",
          density.info=c("none"),
          col=rev(color.palette),
          cexRow=0.7, cexCol=0.75, 
          labCol = FALSE) ##3takes 10 minues
```

## Figure 3B: venn diagrams for # of regions
### Venn diagram data set up
```{r 3b.1, cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE}
# Frequency of regions DA
Freq <- c(nrow(filter(timepoints, DE.Pax_0hpaVSPax_uninjured == "DOWN")),
          nrow(filter(timepoints, DE.Pax_0hpaVSPax_uninjured == "UP")),
          nrow(filter(timepoints, DE.Pax_0hpaVSPax_6hpa == "UP")), 
          nrow(filter(timepoints, DE.Pax_0hpaVSPax_6hpa == "DOWN")), 
          nrow(filter(timepoints, DE.Pax_24hpaVSPax_6hpa == "UP")), 
          nrow(filter(timepoints, DE.Pax_24hpaVSPax_6hpa == "DOWN")), 
          nrow(filter(timepoints, DE.Pax_24hpaVSPax_72hpa == "UP")),
          nrow(filter(timepoints, DE.Pax_24hpaVSPax_72hpa == "DOWN")), 
          nrow(filter(timepoints, timepoints$DE.Pax_72hpaVSPax_uninjured == "UP")), 
          nrow(filter(timepoints, timepoints$DE.Pax_72hpaVSPax_uninjured == "DOWN")),
          nrow(filter(timepoints, DE.All_Tissue_0hpaVSAll_Tissue_uninjured == "DOWN")),
          nrow(filter(timepoints, DE.All_Tissue_0hpaVSAll_Tissue_uninjured == "UP")),
          nrow(filter(timepoints, DE.All_Tissue_0hpaVSAll_Tissue_6hpa == "UP")), 
          nrow(filter(timepoints, DE.All_Tissue_0hpaVSAll_Tissue_6hpa == "DOWN")), 
          nrow(filter(timepoints, DE.All_Tissue_24hpaVSAll_Tissue_6hpa == "UP")), 
          nrow(filter(timepoints, DE.All_Tissue_24hpaVSAll_Tissue_6hpa == "DOWN")), 
          nrow(filter(timepoints, DE.All_Tissue_24hpaVSAll_Tissue_72hpa == "UP")),
          nrow(filter(timepoints, DE.All_Tissue_24hpaVSAll_Tissue_72hpa == "DOWN")), 
          nrow(filter(timepoints, timepoints$DE.All_Tissue_72hpaVSAll_Tissue_uninjured == "UP")), 
          nrow(filter(timepoints, timepoints$DE.All_Tissue_72hpaVSAll_Tissue_uninjured =="DOWN")),
          c(length(unique(timepoints[timepoints$DE.Pax_0hpaVSPax_uninjured == "DOWN",86])),
          length(unique(timepoints[timepoints$DE.Pax_0hpaVSPax_uninjured == "UP",86])),
          length(unique(timepoints[timepoints$DE.Pax_0hpaVSPax_6hpa == "UP",86])), 
          length(unique(timepoints[timepoints$DE.Pax_0hpaVSPax_6hpa == "DOWN",86])), 
          length(unique(timepoints[timepoints$DE.Pax_24hpaVSPax_6hpa == "UP",86])), 
          length(unique(timepoints[timepoints$DE.Pax_24hpaVSPax_6hpa == "DOWN",86])), 
          length(unique(timepoints[timepoints$DE.Pax_24hpaVSPax_72hpa == "UP",86])),
          length(unique(timepoints[timepoints$DE.Pax_24hpaVSPax_72hpa == "DOWN",86])), 
          length(unique(timepoints[timepoints$DE.Pax_72hpaVSPax_uninjured == "UP",86])), 
          length(unique(timepoints[timepoints$DE.Pax_72hpaVSPax_uninjured == "DOWN",86])),
          length(unique(timepoints[timepoints$DE.All_Tissue_0hpaVSAll_Tissue_uninjured == "DOWN",86])),
          length(unique(timepoints[timepoints$DE.All_Tissue_0hpaVSAll_Tissue_uninjured == "UP",86])),
          length(unique(timepoints[timepoints$DE.All_Tissue_0hpaVSAll_Tissue_6hpa == "UP",86])), 
          length(unique(timepoints[timepoints$DE.All_Tissue_0hpaVSAll_Tissue_6hpa == "DOWN",86])), 
          length(unique(timepoints[timepoints$DE.All_Tissue_24hpaVSAll_Tissue_6hpa == "UP",86])), 
          length(unique(timepoints[timepoints$DE.All_Tissue_24hpaVSAll_Tissue_6hpa == "DOWN",86])), 
          length(unique(timepoints[timepoints$DE.All_Tissue_24hpaVSAll_Tissue_72hpa == "UP",86])),
          length(unique(timepoints[timepoints$DE.All_Tissue_24hpaVSAll_Tissue_72hpa == "DOWN",86])), 
          length(unique(timepoints[timepoints$DE.All_Tissue_72hpaVSAll_Tissue_uninjured == "UP",86])), 
          length(unique(timepoints[timepoints$DE.All_Tissue_72hpaVSAll_Tissue_uninjured =="DOWN",86]))))


# Metadata
Timepoint <- c(rep(rep(c("uninjured","0hpa","0hpa","6hpa", "24hpa", "6hpa", "24hpa", "72hpa", "72hpa", "uninjured"), 2),2))
Contrast <- c(rep(rep(c("0vWht", "0vWht", "0v6", "0v6", "24v6", "24v6", "24v72", "24v72", "72vWht", "72vWht"), 2),2))
Condition <- rep(c(rep("pax6", 10), rep("all-tissue", 10)),2)
Direction <- c(rep(rep(c("lose","gain", "lose", "gain", "gain", "lose", "lose", "gain", "lose", "gain"), 2),2))
Measure <- c(rep("peaks",20), rep("genes",20))
peakbreak <- data.frame(Timepoint, Contrast, Condition, Freq, Measure, Direction)
#peakbreak <- peakbreak[order(peakbreak$Contrast),]
# adjust gain direction for pyramid plot
peakbreak$Freq <- ifelse(peakbreak$Direction %in% c("pax6","gain"), -1*peakbreak$Freq, peakbreak$Freq)

# factor for plot ordering
peakbreak$Condition <- factor(peakbreak$Condition, levels = c("pax6", "all-tissue"))
peakbreak$Timepoint <- factor(peakbreak$Timepoint, levels = c("uninjured", "0hpa", "6hpa", "24hpa", "72hpa"))
peakbreak$Direction <- factor(peakbreak$Direction, levels = c("gain", "lose"))

# view data table
datatable(peakbreak, caption = "Break down of peaks for timepoints within a condition")

```

### 6hpa venn diagram
```{r 3b.2, cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE}

grid.newpage()
# values from venn diagram from table
draw.pairwise.venn(area1 = 753,
                   area2 = 1350,
                   cross.area = 170,
                   category = c("0v6hpa", "24v6hpa"),
                   cat.pos = c(0, 0),
                   fill = c("#F7AA14", "#172869"),
                   alpha = c(0.5, 0.5)
                   )

```

### 24hpa venn diagram
```{r 3b.3, cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE}

grid.newpage()
draw.pairwise.venn(area1 = 1758,
                   area2 = 1208,
                   cross.area = 191,
                   category = c("24v6hpa", "24v72hpa"),
                   cat.pos = c(0, 0),
                  fill = c("#F7AA14", "#172869"),
                  alpha = c(0.5, 0.5))

```

### 24hpa venn diagram
```{r 3b.4, cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE}

grid.newpage()
draw.pairwise.venn(area1 = 1002,
                   area2 = 803,
                   cross.area = 118,
                   category = c("24v72hpa", "72vWht"),
                   cat.pos = c(0, 0), 
                   fill = c("#F7AA14", "#172869"),
                   alpha = c(0.5, 0.5)
                   )

```



## Figure 3C: Gene ontology analysis
### 6hpa
#### GO
```{r 3c.1 6hpa, cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE}
# input tss
tss_6hpa_1_Peaks <- rownames(timepoints[timepoints$DE.Pax_0hpaVSPax_6hpa == "DOWN",])
tss_6hpa_2_Peaks <- rownames(timepoints[timepoints$DE.Pax_24hpaVSPax_6hpa == "DOWN",])
tss_6hpa_Peaks <- unique(c(tss_6hpa_1_Peaks, tss_6hpa_2_Peaks))
tss_6hpa <- timepoints[tss_6hpa_Peaks,]
GO_6hpa <- as.data.frame(tss_6hpa$Gene)

input_6hpa <- tss_6hpa[,c(23, 26, 86)]

# view inputs
datatable(input_6hpa, caption = "Input DA regions to GO")

#GO BP 
GO_6hpa_terms <- gost(query = as.character(input_6hpa$Gene), 
                    organism = "hsapiens",
                    sources = "GO:BP",
                    evcodes = TRUE)  ##input gene list, run GO

# Make GO Table
Go_6hpa_use <- GO2table(GO_6hpa_terms)

# View GO table
datatable(Go_6hpa_use)

# Save out GO
write.table(Go_6hpa_use, "~/Desktop/pax6 paper/Final Markdowns/ATAC/ATAC_DA/outs_figures/GO6hpa.txt", sep = "\t")


```
#### ReviGO
```{r 3c.2 6hpa, cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE}

 print(Go_6hpa_use[,c(2,1)])
# Place in Revigo (http://revigo.irb.hr/)
# Allowed SImilarity: small (0.5)
# p-values
# database with GO term sizes: Homo sapiens
# Semantic similarity measure: SimRel
# Save out first page as .csv

# Read in .csv
revi_6hpa <- read.csv("~/Desktop/pax6 paper/Final Markdowns/ATAC/ATAC_DA/outs_figures/revigo_6hpa.csv") # .csv
revi_6hpa <- revi_6hpa[-1*grep(revi_6hpa$plot_X, pattern ="null"),] # remove nested GO terms
revi_6hpa$log10.p.value <- -1*revi_6hpa$log10.p.value
revi_6hpa$description <- tolower(revi_6hpa$description)
# order for barplot
revi_6hpa <- revi_6hpa[order(-revi_6hpa$log10.p.value),]
TermLevels <- concat.GO((revi_6hpa[, 2]))
revi_6hpa$description <- factor(revi_6hpa$description, levels = c(TermLevels))
revi_6hpa <- merge.data.frame(Go_6hpa_use, revi_6hpa, by.x = "term_id", by.y = "term_ID")

# view revigo output
datatable(revi_6hpa)

# plot main GO terms
ggplot(revi_6hpa, aes(x= description, y= log10.p.value))  +
  geom_bar(stat = "identity", fill = "green4") +
  coord_flip()+
  labs(x = "GO Terms", y = "-log(pvalue)", title = "6hpa GO Terms") +
  blank_theme +
  theme(text = element_text(size=20))


# Allowed SImilarity: tiny (0.4)
# p-values
# database with GO term sizes: Homo sapiens
# Semantic similarity measure: SimRel
# Save out first page as .csv

# Read in .csv
revi_6hpa <- read.csv("~/Desktop/pax6 paper/Final Markdowns/ATAC/ATAC_DA/outs_figures/revigo_6hpa_tiny.csv") # .csv
revi_6hpa <- revi_6hpa[-1*grep(revi_6hpa$plot_X, pattern ="null"),] # remove nested GO terms
revi_6hpa$log10.p.value <- -1*revi_6hpa$log10.p.value
revi_6hpa$description <- tolower(revi_6hpa$description)
# order for barplot
revi_6hpa <- revi_6hpa[order(-revi_6hpa$log10.p.value),]
TermLevels <- concat.GO((revi_6hpa[, 2]))
revi_6hpa$description <- factor(revi_6hpa$description, levels = c(TermLevels))
revi_6hpa <- merge.data.frame(Go_6hpa_use, revi_6hpa, by.x = "term_id", by.y = "term_ID")

# view revigo output
datatable(revi_6hpa)

# plot main GO terms
ggplot(revi_6hpa, aes(x= description, y= log10.p.value))  +
  geom_bar(stat = "identity", fill = "green4") +
  coord_flip()+
  labs(x = "GO Terms", y = "-log(pvalue)", title = "6hpa GO Terms") +
  blank_theme +
  theme(text = element_text(size=20))


## Tree map
# Place in Revigo (http://revigo.irb.hr/)
# Allowed SImilarity: small (0.5)
# p-values
# database with GO term sizes: Homo sapiens
# Semantic similarity measure: SimRel
# pin "nephron tubule morphogenesis"
# Save out treemap as .csv
revi_tree_6hpa <- read.csv("~/Desktop/pax6 paper/Final Markdowns/ATAC/ATAC_DA/outs_figures/REVIGO_treemap_6hpa_091319.csv", header = 1)
treegraph(dtf = revi_tree_6hpa, 
          index = c("representative", "description"), show.labels = TRUE)

treegraph(dtf = revi_tree_6hpa, 
          index = c("representative", "description"), 
          show.labels = FALSE, 
          vertex.size = 10,
          rootlabel = "6hpa")
          #vertex.layout="fruchterman.reingold")


## Heatmap of nephron tubule morphogenesis
# define genes 
# combine treemap and go table
revi_heatmap_6 <- merge.data.frame(Go_6hpa_use, revi_tree_6hpa, by.x = "term_id", by.y = "term_ID" )
gene <- concat.GO(filter(revi_heatmap_6, term_name == "nephron tubule morphogenesis")[,4])
gene <- concat.grep(gene)
GOGenes <- grep(gene, tss_6hpa$Gene, ignore.case = T) # get gene name rows from table

# Create heatmap matrix from counts data and input genes
paxGoGenes <- tss_6hpa[GOGenes, ]
paxGoGenes.cpm <- log.counts.timepoints[rownames(paxGoGenes),]
rownames(paxGoGenes.cpm) <- paxGoGenes$Gene
paxGoGenes.cpm <- data.matrix(paxGoGenes.cpm)
paxGoGenes.cpm <- t(scale(t(paxGoGenes.cpm), scale = T))

colnames(paxGoGenes.cpm) <- gsub("_[0-9]$", "", colnames(paxGoGenes.cpm))
colnames(paxGoGenes.cpm) <- gsub("_2.1$", "", colnames(paxGoGenes.cpm))

mcounts <- t(apply(paxGoGenes.cpm, 1, function(x) {
  tapply(x, factor(colnames(paxGoGenes.cpm)), mean)
}))

# colnames(mcounts) # view colnames to pick columns and order for heatmap
mcounts <- (mcounts[,c(6, 8, 7, 9 , 10)])

# Plot heatmap     
heatmap.2((mcounts), 
          scale="none",
          trace="none",
          Colv = FALSE,
          Rowv = TRUE, 
          dendrogram = 'none',
          density.info=c("none"),
          col=rev(color.palette),
          key = FALSE,
          cexRow=2, 
          cexCol=1, 
          labCol = FALSE, margins = c(0,10), ) 

```


### 24hpa
#### GO
```{r 3c.1 24hpa, cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE}

# Input to GO
tss_24hpa_1_Peaks <- rownames(timepoints[timepoints$DE.Pax_24hpaVSPax_6hpa == "UP",])
tss_24hpa_2_Peaks <- rownames(timepoints[timepoints$DE.Pax_24hpaVSPax_72hpa == "UP",])
tss_24hpa_Peaks <- unique(c(tss_24hpa_2_Peaks))
tss_24hpa <- timepoints[tss_24hpa_Peaks,]
GO_24hpa <- as.data.frame(tss_24hpa$Gene)

input_24hpa <- tss_24hpa[,c(26, 27, 86)]

# View inputs
datatable(input_24hpa, caption = "24hpa DA region GO input")

#GO 
GO_24hpa_terms <- gost(query = as.character(input_24hpa$Gene), 
                    organism = "hsapiens",
                    sources = "GO:BP",
                    evcodes = TRUE)  ##input gene list, run GO

# Make GO Table
Go_24hpa_use <- GO2table(GO_24hpa_terms)

# View GO table
datatable(Go_24hpa_use)

# Save out GO
write.table(Go_24hpa_use, "~/Desktop/pax6 paper/Final Markdowns/ATAC/ATAC_DA/outs_figures/GO24hpa.txt", sep = "\t")

```
#### ReviGO
```{r 3c.2 24hpa, cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE}

print(Go_24hpa_use[,c(2,1)])
# Place in Revigo (http://revigo.irb.hr/)
# Allowed SImilarity: small (0.5)
# p-values
# database with GO term sizes: Homo sapiens
# Semantic similarity measure: SimRel
# Save out first page as .csv

# Read in .csv
revi_24hpa <- read.csv("outs_figures/revigo_24hpa.csv") # .csv
revi_24hpa <- revi_24hpa[-1*grep(revi_24hpa$plot_X, pattern ="null"),] # remove nested GO terms
revi_24hpa$log10.p.value <- -1*revi_24hpa$log10.p.value
# order for barplot
revi_24hpa <- revi_24hpa[order(-revi_24hpa$log10.p.value),]
TermLevels <- as.character((revi_24hpa[, 2]))
revi_24hpa$description <- factor(revi_24hpa$description, levels = c(TermLevels))
revi_24hpa <- merge.data.frame(Go_24hpa_use, revi_24hpa, by.x = "term_id", by.y = "term_ID")

# view revigo output
datatable(revi_24hpa)

# plot main GO terms
ggplot(revi_24hpa, aes(x= description, y= log10.p.value))  +
  geom_bar(stat = "identity", fill = "green4") +
  coord_flip()+
  labs(x = "GO Terms", y = "-log(pvalue)", title = "24hpa GO Terms") +
  blank_theme+
  theme(text = element_text(size=20))



# Place in Revigo (http://revigo.irb.hr/)
# Allowed SImilarity: tiny (0.4)
# p-values
# database with GO term sizes: Homo sapiens
# Semantic similarity measure: SimRel
# Save out first page as .csv

# Read in .csv
revi_24hpa <- read.csv("outs_figures/revigo_24hpa_tiny.csv") # .csv
revi_24hpa <- revi_24hpa[-1*grep(revi_24hpa$plot_X, pattern ="null"),] # remove nested GO terms
revi_24hpa$log10.p.value <- -1*revi_24hpa$log10.p.value
# order for barplot
revi_24hpa <- revi_24hpa[order(-revi_24hpa$log10.p.value),]
TermLevels <- as.character((revi_24hpa[, 2]))
revi_24hpa$description <- factor(revi_24hpa$description, levels = c(TermLevels))
revi246hpa <- merge.data.frame(Go_24hpa_use, revi_24hpa, by.x = "term_id", by.y = "term_ID")

# view revigo output
datatable(revi_24hpa)

# plot main GO terms
ggplot(revi_24hpa, aes(x= description, y= log10.p.value))  +
  geom_bar(stat = "identity", fill = "green4") +
  coord_flip()+
  labs(x = "GO Terms", y = "-log(pvalue)", title = "24hpa GO Terms") +
  blank_theme+
  theme(text = element_text(size=20))


## Tree map
# Place in Revigo (http://revigo.irb.hr/)
# Allowed SImilarity: small (0.5)
# p-values
# database with GO term sizes: Homo sapiens
# Semantic similarity measure: SimRel
# pin "neuron development"
# Save out treemap as .csv

revi_tree_24hpa <- read.csv("outs_figures/REVIGO_treemap_24hpa_091319.csv", header = 1)
treegraph(dtf = revi_tree_24hpa, 
          index = c("representative", "description"), show.labels = TRUE)
treegraph(dtf = revi_tree_24hpa, 
          index = c("representative", "description"), 
          show.labels = FALSE,
          vertex.size = 10,
          rootlabel = "24hpa")


## Heatmap of neuron development
# define genes 
# combine treemap and go table
revi_heatmap_24 <- merge.data.frame(Go_24hpa_use, revi_tree_24hpa, by.x = "term_id", by.y = "term_ID" )
gene <- concat.GO(filter(revi_heatmap_24, term_name == "neuron development")[,4])
gene <- concat.grep(gene)
GOGenes <- grep(gene, tss_24hpa$Gene, ignore.case = T) # get gene name rows from table

# Create heatmap matrix from counts data and input genes
paxGoGenes <- tss_24hpa[GOGenes[1:20], ]
paxGoGenes.cpm <- log.counts.timepoints[rownames(paxGoGenes),]
rownames(paxGoGenes.cpm) <- paxGoGenes$Gene
paxGoGenes.cpm <- data.matrix(paxGoGenes.cpm)
paxGoGenes.cpm <- t(scale(t(paxGoGenes.cpm), scale = T))

colnames(paxGoGenes.cpm) <- gsub("_[0-9]$", "", colnames(paxGoGenes.cpm))
colnames(paxGoGenes.cpm) <- gsub("_2.1$", "", colnames(paxGoGenes.cpm))

mcounts <- t(apply(paxGoGenes.cpm, 1, function(x) {
  tapply(x, factor(colnames(paxGoGenes.cpm)), mean)
}))

# colnames(mcounts) # view colnames to pick columns and order for heatmap
mcounts <- (mcounts[,c(6, 8, 7, 9 , 10)])

# Plot heatmap     
heatmap.2((mcounts), 
          scale="none",
          trace="none",
          Colv = FALSE,
          Rowv = TRUE, 
          dendrogram = 'none',
          density.info=c("none"),
          col=rev(color.palette),
          key = FALSE,
          cexRow=2, 
          cexCol=1, 
          labCol = FALSE, margins = c(0,10))

```

### 72hpa
#### GO
```{r 3c.1 72hpa, cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE}

# Inputs
tss_72hpa_1_Peaks <- rownames(timepoints[timepoints$DE.Pax_24hpaVSPax_72hpa == "DOWN",])
tss_72hpa_2_Peaks <- rownames(timepoints[timepoints$DE.Pax_72hpaVSPax_Whole.Tail == "UP",])
tss_72hpa_Peaks <- unique(c(tss_72hpa_1_Peaks, tss_72hpa_2_Peaks))
tss_72hpa <- timepoints[tss_72hpa_Peaks,]
GO_72hpa <- as.data.frame(tss_72hpa$Gene)


input_72hpa <- tss_72hpa[,c(27, 31, 86)]

# view inputs
datatable(input_72hpa, caption = "72hpa DA region GO input")

#GO BP#
GO_72hpa_terms <- gost(query = as.character(input_72hpa$Gene), 
                    organism = "hsapiens",
                    sources = "GO:BP",
                    evcodes = TRUE)  ##input gene list, run GO

# Make GO Table
Go_72hpa_use <- GO2table(GO_72hpa_terms)

# View GO table
datatable(Go_72hpa_use)

# Save out GO
write.table(Go_72hpa_use, "~/Desktop/pax6 paper/Final Markdowns/ATAC/ATAC_DA/outs_figures/GO72hpa.txt", sep = "\t")

```

#### ReviGO
```{r 3c.2 72hpa, cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE}

print(Go_72hpa_use[,c(2,1)])
# Place in Revigo (http://revigo.irb.hr/)
# Allowed SImilarity: small (0.5)
# p-values
# database with GO term sizes: Homo sapiens
# Semantic similarity measure: SimRel
# Save out first page as .csv

# Read in .csv
revi_72hpa <- read.csv("outs_figures/revigo_72hpa.csv") # .csv
revi_72hpa <- revi_72hpa[-1*grep(revi_72hpa$plot_X, pattern ="null"),] # remove nested GO terms
revi_72hpa$log10.p.value <- -1*revi_72hpa$log10.p.value
# order for barplot
revi_72hpa <- revi_72hpa[order(-revi_72hpa$log10.p.value),]
TermLevels <- as.character((revi_72hpa[, 2]))
revi_72hpa$description <- factor(revi_72hpa$description, levels = c(TermLevels))
revi_72hpa <- merge.data.frame(Go_72hpa_use, revi_72hpa, by.x = "term_id", by.y = "term_ID")

# view revigo output
datatable(revi_72hpa)

# plot main GO terms
ggplot(revi_72hpa, aes(x= description, y= log10.p.value))  +
  geom_bar(stat = "identity", fill = "green4") +
  coord_flip()+
  labs(x = "GO Terms", y = "-log(pvalue)", title = "72hpa GO Terms") +
  blank_theme+
  theme(text = element_text(size=20))


# Place in Revigo (http://revigo.irb.hr/)
# Allowed SImilarity: small (0.5)
# p-values
# database with GO term sizes: Homo sapiens
# Semantic similarity measure: SimRel
# Save out first page as .csv

# Read in .csv
revi_72hpa <- read.csv("outs_figures/revigo_72hpa_tiny.csv") # .csv
revi_72hpa <- revi_72hpa[-1*grep(revi_72hpa$plot_X, pattern ="null"),] # remove nested GO terms
revi_72hpa$log10.p.value <- -1*revi_72hpa$log10.p.value
# order for barplot
revi_72hpa <- revi_72hpa[order(-revi_72hpa$log10.p.value),]
TermLevels <- as.character((revi_72hpa[, 2]))
revi_72hpa$description <- factor(revi_72hpa$description, levels = c(TermLevels))
revi_72hpa <- merge.data.frame(Go_72hpa_use, revi_72hpa, by.x = "term_id", by.y = "term_ID")

# view revigo output
datatable(revi_72hpa)

# plot main GO terms
ggplot(revi_72hpa, aes(x= description, y= log10.p.value))  +
  geom_bar(stat = "identity", fill = "green4") +
  coord_flip()+
  labs(x = "GO Terms", y = "-log(pvalue)", title = "72hpa GO Terms") +
  blank_theme+
  theme(text = element_text(size=25))


## Tree map
# Place in Revigo (http://revigo.irb.hr/)
# Allowed SImilarity: small (0.5)
# p-values
# database with GO term sizes: Homo sapiens
# Semantic similarity measure: SimRel
# Save out treemap as .csv

revi_tree_72hpa <- read.csv("outs_figures/REVIGO_treemap_72hpa_091319.csv", header = 1)
treegraph(dtf = revi_tree_72hpa, 
          index = c("representative", "description"), show.labels = TRUE)
treegraph(dtf = revi_tree_72hpa, 
          index = c("representative", "description"), 
          show.labels = FALSE, 
          vertex.size = 10,
          rootlabel = "72hpa")


## Heatmap of neuron development
# define genes 
# combine treemap and go table
revi_heatmap_72 <- merge.data.frame(Go_72hpa_use, revi_tree_72hpa, by.x = "term_id", by.y = "term_ID" )
gene <- concat.GO(filter(revi_heatmap_72, term_name == "neural precursor cell proliferation")[,4])
gene <- concat.grep(gene)
GOGenes <- grep(gene, tss_72hpa$Gene, ignore.case = T) # get gene name rows from table

# Create heatmap matrix from counts data and input genes
paxGoGenes <- tss_72hpa[GOGenes, ]
paxGoGenes.cpm <- log.counts.timepoints[rownames(paxGoGenes),]
rownames(paxGoGenes.cpm) <- paxGoGenes$Gene
paxGoGenes.cpm <- data.matrix(paxGoGenes.cpm)
paxGoGenes.cpm <- t(scale(t(paxGoGenes.cpm), scale = T))

colnames(paxGoGenes.cpm) <- gsub("_[0-9]$", "", colnames(paxGoGenes.cpm))
colnames(paxGoGenes.cpm) <- gsub("_2.1$", "", colnames(paxGoGenes.cpm))

mcounts <- t(apply(paxGoGenes.cpm, 1, function(x) {
  tapply(x, factor(colnames(paxGoGenes.cpm)), mean)
}))

# colnames(mcounts) # view colnames to pick columns and order for heatmap
mcounts <- (mcounts[,c(6, 8, 7, 9 , 10)])

# Plot heatmap     
heatmap.2((mcounts), 
          scale="none",
          trace="none",
          Colv = FALSE,
          Rowv = TRUE, 
          dendrogram = 'none',
          density.info=c("none"),
          col=rev(color.palette),
          key = FALSE,
          cexRow=2, 
          cexCol=1, 
          labCol = FALSE, margins = c(0,10))


```


## 5C/D: pbx3 and meis1 target heatmaps
### Read in tables if not continuing from above
```{r grn tables, cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE}

# read in DA motif tables from GRN prediction
table.6 <- read.table("~/Desktop/pax6 paper/Final Markdowns/GRN/outs/table.6.txt", sep = "\t")
table.24 <- read.table("~/Desktop/pax6 paper/Final Markdowns/GRN/outs/table.24.txt", sep = "\t")
table.72 <- read.table("~/Desktop/pax6 paper/Final Markdowns/GRN/outs/table.72.txt", sep = "\t")

# make master list of tables
mastertable <- rbind(table.24, table.6, table.72)


```

### Identify other inputs and outs of pbx3 and meis 1
```{r pbx3 and meis1, cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE}

# # filter master table for rows where TF is pbx3 and the gene is pbx3
# Master_pbx <- filter(mastertable, TF == "pbx3" | Gene =="pbx3")
# # filter master table for rows where TF is meis1 and the gene is meis1
# Master_meis <- filter(mastertable, TF == "meis1" | Gene =="meis1")

#Master_pbx_meis <- rbind(Master_pbx, Master_meis)

# filter master table for rows where TF is pbx3 
Master_pbx <- filter(mastertable, TF == "pbx3")
# filter master table for rows where TF is meis1 
Master_meis <- filter(mastertable, TF == "meis1")
Master_pbx_meis <- filter(mastertable, TF == "pbx3" | TF == "meis1")


# view table
datatable(Master_pbx_meis)


```

### pbx3 targets 1
```{r pbx3 targets 1, cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE}
# Identify genes that pbx3 targets at each timepoint
# 6hpa
treegraph(filter(Master_pbx, timepoint=="6hpa" & TF == "pbx3"), index = c("TF", "Gene"), directed = TRUE, show.labels = TRUE, rootlabel = "6hpa pbx3 targets")
pbx3_6 <- as.character(filter(Master_pbx, timepoint=="6hpa")[,1]) # target peaks
pbx3_6_gene <- as.character(filter(Master_pbx, timepoint=="6hpa")[,17]) # target genes


# 24hpa
treegraph(filter(Master_pbx, timepoint=="24hpa" & TF == "pbx3"), index = c("TF", "Gene"), directed = TRUE, show.labels = TRUE, rootlabel = "24hpa pbx3 targets")
pbx3_24 <- as.character(filter(Master_pbx, timepoint=="24hpa")[,1]) # target peaks
pbx3_24_gene <- as.character(filter(Master_pbx, timepoint=="24hpa")[,17]) # target genes

 

# 72hpa
treegraph(filter(Master_pbx, timepoint=="72hpa" & TF == "pbx3"), index = c("TF", "Gene"), directed = TRUE, show.labels = TRUE, rootlabel = "72hpa pbx3 targets")
pbx3_72 <- as.character(filter(Master_pbx, timepoint=="72hpa")[,1]) # target peaks
pbx3_72_gene <- as.character(filter(Master_pbx, timepoint=="72hpa")[,17]) # target genes


pbx_all <- unique(c(pbx3_6, pbx3_24, pbx3_72))
pbx_all_gene <- unique(c(pbx3_6_gene, pbx3_24_gene, pbx3_72_gene))
# Make heatmap for pbx3 targets
gene <- concat.grep(pbx3_72)
GOGenes <- grep(pattern = gene, timepoints$Peak, ignore.case = T) # get gene name rows from table

# Create heatmap matrix from counts data and input genes
paxGoGenes <- timepoints[GOGenes, ]
paxGoGenes <- paxGoGenes[-grep(paxGoGenes$Gene, pattern = "Xetrov|LOC", ignore.case = TRUE),] # remove unannotated
paxGoGenes.cpm <- log.counts.timepoints[rownames(paxGoGenes),]
rownames(paxGoGenes.cpm) <- paxGoGenes$Gene
paxGoGenes.cpm <- data.matrix(paxGoGenes.cpm)
paxGoGenes.cpm <- t(scale(t(paxGoGenes.cpm), scale = T))

colnames(paxGoGenes.cpm) <- gsub("_[0-9]$", "", colnames(paxGoGenes.cpm))
colnames(paxGoGenes.cpm) <- gsub("_2.1$", "", colnames(paxGoGenes.cpm))

mcounts <- t(apply(paxGoGenes.cpm, 1, function(x) {
  tapply(x, factor(colnames(paxGoGenes.cpm)), mean)
}))

#colnames(mcounts) # view colnames to pick columns and order for heatmap
mcounts <- (mcounts[,c(6, 8, 7, 9 , 10)])

# Plot heatmap     
heatmap.2((mcounts), 
          scale="none",
          trace="none",
          Colv = FALSE,
          Rowv = TRUE, 
          dendrogram = 'none',
          density.info=c("none"),
          col=rev(color.palette),
          key = FALSE,
          cexRow=2, 
          cexCol=1, 
          labCol = FALSE, margins = c(0,10)) 



```


### meis1 targets
```{r meis targets, cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE}
# Identify genes that meis1 targets at each timepoint
# 6hpa
treegraph(filter(Master_meis, timepoint=="6hpa" & TF == "meis1"), index = c("TF", "Gene"), directed = TRUE, show.labels = TRUE, rootlabel = "6hpa meis1 targets")
meis1_6 <- as.character(filter(Master_meis, timepoint=="6hpa")[,1]) # target peaks
meis1_6_gene <- as.character(filter(Master_meis, timepoint=="6hpa")[,17]) # target genes

# 24hpa
treegraph(filter(Master_meis, timepoint=="24hpa" & TF == "meis1"), index = c("TF", "Gene"), directed = TRUE, show.labels = TRUE, rootlabel = "24hpa meis1 targets")
meis1_24 <- as.character(filter(Master_meis, timepoint=="24hpa")[,1]) # target peaks
meis1_24_gene <- as.character(filter(Master_meis, timepoint=="24hpa")[,17]) # target genes


# 72hpa
treegraph(filter(Master_meis, timepoint=="72hpa" & TF == "meis1"), index = c("TF", "Gene"), directed = TRUE, show.labels = TRUE, rootlabel = "72hpa meis1 targets")
meis1_72 <- as.character(filter(Master_meis, timepoint=="72hpa")[,1]) #target peaks
meis1_72_gene <- as.character(filter(Master_meis, timepoint=="72hpa")[,17]) # target genes



meis_all <- unique(c(meis1_6, meis1_24, meis1_72)) # target peaks
# Make heatmap for pbx3 targets
gene <- concat.grep(meis1_24)
GOGenes <- grep(gene, timepoints$Peak, ignore.case = T) # get gene name rows from table

# Create heatmap matrix from counts data and input genes
paxGoGenes <- timepoints[GOGenes, ]
paxGoGenes <- paxGoGenes[-grep(paxGoGenes$Gene, pattern = "Xetrov|LOC|-like", ignore.case = TRUE),] # remove unannotated
paxGoGenes <- paxGoGenes[1:20,]
paxGoGenes.cpm <- log.counts.timepoints[rownames(paxGoGenes),]
rownames(paxGoGenes.cpm) <- paxGoGenes$Gene
paxGoGenes.cpm <- data.matrix(paxGoGenes.cpm)
paxGoGenes.cpm <- t(scale(t(paxGoGenes.cpm), scale = T))

colnames(paxGoGenes.cpm) <- gsub("_[0-9]$", "", colnames(paxGoGenes.cpm))
colnames(paxGoGenes.cpm) <- gsub("_2.1$", "", colnames(paxGoGenes.cpm))

mcounts <- t(apply(paxGoGenes.cpm, 1, function(x) {
  tapply(x, factor(colnames(paxGoGenes.cpm)), mean)
}))

# colnames(mcounts) # view colnames to pick columns and order for heatmap
mcounts <- (mcounts[,c(6, 8, 7, 9 , 10)])

# Plot heatmap     
heatmap.2((mcounts), 
          scale="none",
          trace="none",
          Colv = FALSE,
          Rowv = TRUE, 
          dendrogram = 'none',
          density.info=c("none"),
          col=rev(color.palette),
          key = FALSE,
          cexRow=2, 
          cexCol=1, 
          labCol = FALSE, margins = c(0,10)) 

```

## Figure 5B: peaks with both meis1 and pbx3 sites
### Peaks with both meis1 and pbx3 sites
```{r Figure 5B, cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE}

pbx_all <- unique(c(pbx3_6, pbx3_24, pbx3_72))
meis_all <- unique(c(meis1_6, meis1_24, meis1_72))

length(meis_all)
length(pbx_all)

co6hpa <- intersect(meis_all, pbx_all)
length(co6hpa)

grid.newpage()
draw.pairwise.venn(area1 = 282,
                   area2 = 61,
                   cross.area = 26,
                   cat.pos = c(0, 0),
                   alpha = c(0.5, 0.5), 
                   rotation.degree = 270,
                   fill = c("#B574A7","#FBB58F")
                   )

```