---
title: "IWRvUninj Seurat saturated"
author: "Anneke Kakebeen"
date: "4/9/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Set up
```{r setup, include=FALSE}

# Load Packages
library(Seurat)
library(ggplot2)
library(LaCroixColoR)
library(RColorBrewer)
library(cowplot)
library(DT)
library(dplyr)
library(gridExtra)

## This is where files will save ##
setwd("/Users/annekekakebeen/Desktop/pax6 paper/Final Markdowns/SingleCell_1/analysis/")
source("~/Desktop/pax6 paper/Final Markdowns/common_source_functions.R")

```

## DE functions
```{r DE functions}
# Define DE marker function
condition.marker.1 <- function(object, cluster1){ 
   FindMarkers(object = object, ident.1 = cluster1, verbose = FALSE, min.pct = 0.3, only.pos = TRUE) #cluster = to one cluster denoted by a number
}

# Define function for loop for marker function
# clusters = list of clusters by number, outfiles= names out objects to be made
cluster.marker.multi.1 <- function(object, clusters1, outfiles){ 
  for(i in 1:length(clusters1)){
      # Perform Findmarkers function from Seurat
      temp.markers <- condition.marker.1(object, clusters1[i])
      # Get number of rows in tempprary table
      n <- nrow(temp.markers)
      # Add column with identifier
      temp.markers$cluster <- rep(outfiles[i], n)
      # make index of rows with LOC
      remove <- grep(pattern = "LOC.*", x = rownames(temp.markers))
      # remove rows with LOC
      temp.table <- temp.markers[-remove,]
      # make index to remove rows wtih Xetrov
      remove <- grep(pattern = "Xetrov.*", x = rownames(temp.table))
      # remove rows with Xetrov
      temp.table <- temp.table[-remove,]
      # make column for genes
      temp.table$gene <- rownames(temp.table)
      # Assign to outfile name in the enviornment
    assign(outfiles[i], temp.table, envir = .GlobalEnv)
    }
}


### de genes with two idents
# Define DE marker function
condition.marker.2 <- function(object, cluster1, cluster2){ 
   FindMarkers(object = object, ident.1 = cluster1, ident.2 = cluster2, verbose = FALSE, min.pct = 0.3, only.pos = TRUE, min.cells.group = 0, assay = "RNA") #cluster = to one cluster denoted by a number
}

# Define function for loop for marker function
# clusters = list of clusters by number, outfiles= names out objects to be made
cluster.marker.multi.2 <- function(object, clusters1, clusters2, outfiles){ 
  for(i in 1:length(clusters1)){
    for(x in 1:length(clusters2)){
      # Perform Findmarkers function from Seurat
      temp.markers <- condition.marker.2(object, clusters1[i], clusters2[x])
      # Get number of rows in tempprary table
      n <- nrow(temp.markers)
      # Add column with identifier
      temp.markers$cluster <- rep(outfiles[i], n)
      # make index of rows with LOC
      remove <- grep(pattern = "LOC.*", x = rownames(temp.markers))
      # remove rows with LOC
      temp.table <- temp.markers[-remove,]
      # make index to remove rows wtih Xetrov
      remove <- grep(pattern = "Xetrov.*", x = rownames(temp.table))
      # remove rows with Xetrov
      temp.table <- temp.table[-remove,]
      temp.table$gene <- rownames(temp.table)
      # Assign to outfile name in the enviornment
    assign(outfiles[i], temp.table, envir = .GlobalEnv)
    }}
}
```


## Read in Seurat Objects
```{r load objects, cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE}

# Read in time.neural seurat object
pax.neural <- readRDS("~/Desktop/pax6 paper/Final Markdowns/SingleCell_1/UMAP/pax.neural_FINAL_1.RDS")
pax.all <- readRDS("~/Desktop/pax6 paper/Final Markdowns/SingleCell_1/UMAP/pax.combined_FINAL_1.RDS")

```

## Neural
### DE between clusters
```{r DE between clusters}

# Set identity to AK_Names
Idents(pax.neural) <- "Laevis"

# Define ins
clusters<- levels(pax.neural$Laevis) # define clusters to use
clusters <- sort(clusters) # sort object of cluster names alphabetically

# Outfile names
outfiles <- paste("DEClusters", unique(gsub(clusters, pattern = "_.*", replacement = "")), sep = ".")
outfiles <- sort(outfiles)

# Run function
cluster.marker.multi.1(object =  pax.neural, clusters1 = clusters, outfiles = outfiles)

# Make master table
objects.list <- lapply(outfiles, get)
DEClusters.neural <- do.call(bind_rows, objects.list) # row bind to make large table

# view table
datatable(DEClusters.neural)

# Save out table
write.table(DEClusters.neural, "outs/DEClusters.neural.txt", sep = "\t", col.names = NA)

```

### DE between conditions within one cluster
#### Set up
```{r DE between conditions}

# Define ins
Idents(pax.neural) <- pax.neural$threestate_condition
clusters<- levels(pax.neural@active.ident)

# Uninjured in
clusters_uninj <- clusters[grep(pattern = ".*uninj", x = clusters, value = FALSE)]
clusters_uninj <- sort(clusters_uninj)


# dmso in
clusters_regenerate <- clusters[grep(pattern = ".*regenerate", x = clusters, value = FALSE)]
clusters_regenerate <- sort(clusters_regenerate)

```

#### Lose uninj > 24hpa
```{r DE lose}
# outs
outfiles <- paste("DEConditions_lose", unique(gsub(clusters, pattern = ".*_", replacement = "")), sep = ".")
outfiles <- sort(outfiles)

# run DE analysis
cluster.marker.multi.2(pax.neural, clusters1 = clusters_uninj, clusters2 = clusters_regenerate, outfiles = outfiles)

# make master table
objects.list <- lapply(outfiles, get) # list all outfiles
DEConditions_lose <- do.call(bind_rows, objects.list) # rowbind outfiles into one table
colnames(DEConditions_lose)[3:4] <- c("pct.uninj", "pct.24hpa")

# view table
datatable(DEConditions_lose)

# Save out table
write.table(DEConditions_lose, "outs/DEConditions_lose.txt", sep = "\t", col.names = NA)

```

#### Gain uninj < 24hpa
```{r DE gain}
# outs
outfiles <- paste("DEConditions_gain", unique(gsub(clusters, pattern = ".*_", replacement = "")), sep = ".")
outfiles <- sort(outfiles)

# run DE analysis
cluster.marker.multi.2(pax.neural, clusters1 = clusters_regenerate, clusters2 = clusters_uninj, outfiles = outfiles)

# make master table
objects.list <- lapply(outfiles, get) # list all outfiles
DEConditions_gain <- do.call(bind_rows, objects.list) # rowbind outfiles into one table
colnames(DEConditions_gain)[3:4] <- c("pct.24hpa", "pct.uninj") # rename pct columns for ease

# view table
datatable(DEConditions_gain)

# Save out table
write.table(DEConditions_gain, "outs/DEConditions_gain.txt", sep = "\t", col.names = NA)

```

### DE between conditions when all cells in a condition are treated as an aggregate
#### Set up
```{r DE between clusters agg}

Idents(pax.neural) <- pax.neural.subset$condition

# Define ins
clusters<- levels(pax.neural@active.ident)

```

#### Lose uninj > 24hpa
```{r DE lose agg}
# outs
outfiles <- paste("DEConditions_lose_agg", unique(gsub(clusters, pattern = ".*_", replacement = "")), sep = ".")
outfiles <- sort(outfiles)

# run DE analysis
DEConditions_lose_agg <- condition.marker.2(pax.neural, "uninj", "regenerate")
colnames(DEConditions_lose_agg)[3:4] <- c("pct.uninj", "pct.24hpa")

# view table
datatable(DEConditions_lose_agg)

# Save out table
write.table(DEConditions_lose_agg, "outs/DEConditions_lose_agg.txt", sep = "\t", col.names = NA)

```

#### Gain uninj < 24hpa
```{r DE gain agg}
# outs
outfiles <- paste("DEConditions_gain_agg", unique(gsub(clusters, pattern = ".*_", replacement = "")), sep = ".")
outfiles <- sort(outfiles)

# run DE analysis
DEConditions_gain_agg <- condition.marker.2(pax.neural, "regenerate", "uninj")
colnames(DEConditions_lose_agg)[3:4] <- c("pct.24hpa", "pct.uninj")

# view table
datatable(DEConditions_lose_agg)

# Save out table
write.table(DEConditions_lose_agg, "outs/DEConditions_gain_agg.txt", sep = "\t", col.names = NA)

```



## All cell types
### DE between clusters
```{r DE between clusters all}

# Define ins
Idents(pax.all) <- pax.all$AK_Names
clusters<- levels(pax.all) # define clusters to use

# Outfile names
outfiles <- paste("DEClusters", unique(gsub(clusters, pattern = "_.*", replacement = "")), sep = ".")

# Run function
cluster.marker.multi.1(object =  pax.all, clusters1 = clusters, outfiles = outfiles)

# Make master table
objects.list <- lapply(outfiles, get)
DEClusters.neural <- do.call(bind_rows, objects.list) # row bind to make large table

# view table
datatable(DEClusters.neural)

# Save out table
write.table(DEClusters.neural, "outs/DEClusters.all.txt", sep = "\t", col.names = NA)

```

### Plot of heatmap for top 5 genes in DE
```{r heatmap }

de <- read.table("outs/DEClusters.all.txt", sep = "\t", header = TRUE)
de$diff <- de$pct.1-de$pct.2

genes <- as.data.frame(de %>%
  group_by(cluster) %>%
  top_n(20, diff)) %>%
  group_by(cluster) %>%
  top_n(10, p_val)
datatable(genes)

DoHeatmap(pax.all, features = as.character(genes$gene), assay = "integrated") + NoLegend()
DotPlot(pax.all, features = as.character(unique(genes$gene)), cols = c("blue", "gold")) + RotatedAxis()

```

### All markers between clusters using Seurat "FindAllMarkers" command
```{r all markers in neural and all}

## Get all markers for neural subset
# Idents(pax.neural) <- pax.neural$Laevis
# allneural <- FindAllMarkers(pax.neural, verbose = TRUE)
# allneural$deltapct <- allneural$pct.1 - allneural$pct.2
# allneural <- subset(allneural, deltapct > 0)
# datatable(allneural, caption = "all neural markers")
# 
# insitu <- as.data.frame(allneural %>%
#                           group_by(cluster) %>%
#                           top_n(n = 20, wt = c(deltapct & p_val_adj)))
# 
# datatable(insitu)
# write.table(insitu, "~/Desktop/insitu.txt", sep = "\t")
# 
# # general sense of genes on xenbase
# xenbase <- read.csv("~/Desktop/xenbase.csv")
# 
# # input range of markers table to invesitage
# xenbase_mine <- function(df){
#   table <- merge.data.frame(df, xenbase, by.x = "gene", by.y = "GENE")
#   return(table)
# }
# 
# test <- xenbase_mine(insitu))
# datatable(test)
# 
# 
# ## Get all markers for neural subset
# Idents(pax.all) <- pax.all$AK_Names
# allall$deltapct <- allall$pct.1 - allall$pct.2
# allall <- FindAllMarkers(pax.all, verbose = TRUE)
# datatable(allall, caption = "all  markers")
# 
# 
# ## Get all markers for neural subset
# Idents(pax.all) <- pax.neural$AK_Names
# allall$deltapct <- allall$pct.1 - allall$pct.2
# allall <- FindAllMarkers(pax.all, verbose = TRUE)
# datatable(allall, caption = "all  markers")


```

### De between dmso neuron and uninj neuron
```{r de neuron}

## DE between ocnditions within neuron
Idents(pax.neural) <- pax.neural$threestate_condition
deneuron <- FindMarkers(pax.neural, ident.1 = "regenerate_Neuron", ident.2 = "uninj_Neuron", verbose = TRUE, only.pos = TRUE)
deneuron$deltapct <- deneuron$pct.1 - deneuron$pct.2
deneuron$gene <- rownames(deneuron)

# Save out table
write.table(deneuron, "outs/DEneurons.condition.txt", sep = "\t", col.names = NA)

## DE between neurons and diff and nsc
Idents(pax.neural) <- pax.neural$three_state
deneuron1 <- FindMarkers(pax.neural, ident.1 = "Neuron",verbose = TRUE, only.pos = TRUE)
deneuron1$cluster <- 'neuron'
deneuron2 <- FindMarkers(pax.neural, ident.1 = "NSC",verbose = TRUE, only.pos = TRUE)
deneuron2$cluster <- 'NSC'
deneuron3 <- FindMarkers(pax.neural, ident.1 = "Differentiating",verbose = TRUE, only.pos = TRUE)
deneuron3$cluster <- 'Differentiating'
#combine tables
deneuron <- rbind(deneuron1, deneuron2, deneuron3)
deneuron$deltapct <- deneuron$pct.1 - deneuron$pct.2
deneuron$gene <- rownames(deneuron)

# Save out table
write.table(deneuron, "outs/DEneurons.celltype.txt", sep = "\t", col.names = NA)



```






