---
title: "Single Cell Figures"
author: "Anneke Kakebeen"
date: "13 Aug 2019"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
editor_options: 
  chunk_output_type: console
---

## Set up for scRNA-Seq Figures
```{r scrnaseq figure set up, cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE}
# load packages
library(Seurat)
library(LaCroixColoR)
library(ggplot2)
library(cowplot)
library(tidyr)
library(DT)
library(dplyr)

# load functions
source("~/Desktop/pax6 paper/Final Markdowns/common_source_functions.R")

# Set colors
my.color.neural <- c("#2CB11B", "#F73BEE", "#E9E4A6", "#DFCEE0" , "#E9A17C", "#1BB6AF" , "#172869" , "#FF3200") 
lacroix_palette("PassionFruit"), lacroix_palette("Tangerine"),  lacroix_palette("PinaFraise")))
condition.color <- c( "#6F0909", "#C0C0C0")
cc_colors <- c((lacroix_palette("Lime"))[c(1,3)], "black")

# Set working directory
setwd("~/Desktop/pax6 paper/Final Markdowns/SingleCell_1/analysis/")

```



## Read in scRNASeq Objects
```{r load objects, cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE, cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE}

# Read in pax.all seurat object
pax.all <- readRDS("~/Desktop/pax6 paper/Final Markdowns/SingleCell_1/UMAP/pax.combined_FINAL_1.RDS")

# read in pax.neural
pax.neural <- readRDS("~/Desktop/pax6 paper/Final Markdowns/SingleCell_1/UMAP/pax.neural_FINAL_1.RDS")

# read in plotting data
pax.all.plotting<- read.table("~/Desktop/pax6 paper/Final Markdowns/SingleCell_1/UMAP/pax.combined_metadata.txt", sep = "\t")

# factor plotting table for plotting
pax.all.plotting$condition <- factor(pax.all.plotting$condition, levels = c("uninj", "regenerate"))
pax.all.plotting$Phase <- factor(pax.all.plotting$Phase, levels = c("S", "G2M", "G1"))
```


## Supp 4A: UMAP of all cells colored by individual cell types
```{r supp 4a, cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE}

# change identity to cell clsuter
Idents(pax.all) <- pax.all$AK_Names

# plot cells in umap
DimPlot(pax.all, cols = my.color.neural) +
  NoAxes() +
  NoLegend()

```

## Supp 4B: featureplot for neural cell markers
```{r supp 4c, cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE}

# set assay
DefaultAssay(pax.all) <- "RNA"
# plot neural markers

p1 <- FeaturePlot(pax.all, features = c("sox2"),  order = TRUE, cols = c("gold", "green", "purple"))+ NoAxes()+NoLegend()
p2 <- FeaturePlot(pax.all, features = c("neurog1"),  order = TRUE, cols = c("gold", "green", "purple"))+ NoAxes()+NoLegend()
p3 <- FeaturePlot(pax.all, features = c("neurod1"),  order = TRUE, cols = c("gold", "green", "purple"))+ NoAxes()+NoLegend()
p4 <- FeaturePlot(pax.all, features = c("elavl4"),  order = TRUE, cols = c("gold", "green", "purple"))+ NoAxes()+NoLegend()

plot_grid(p1, p2, p3, p4, nrow = 2)

```

## Supp 4C: UMAP of all cells colored by cell classes
```{r supp 4b, cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE}

Idents(pax.all) <- pax.all$CellClass
DimPlot(pax.all, label = FALSE, cols = my.color.neural[c(1, 3, 5, 7, 9, 11)]) +
  NoAxes() +
  NoLegend()


```



## Supp 4D: Average expression bar plots of pbx3 and meis1 in all cells

### violin plots
```{r vln pbx3 meis}
DefaultAssay(pax.all) <- "RNA"
### vln plot of meis1 and pbx3 from all
all.expression <- FetchData(pax.all, c("ident", "AK_Names","CellClass", "condition", "pbx3", "meis1"))

ggplot(all.expression, aes(x=CellClass, y=pbx3, fill=CellClass)) +
  geom_violin() +
  geom_jitter(size=0.3) +
  theme_bw(base_size = 30) +
  theme(legend.position = "none", axis.text = element_text(colour = "black")) +
    scale_fill_manual(values = my.color.neural[c(1, 3, 5, 7, 9, 11)]) +
  xlab("")

ggplot(all.expression, aes(x=CellClass, y=meis1, fill=CellClass)) +
  geom_violin() +
  geom_jitter(size=0.3) +
  theme_bw(base_size = 30) +
  scale_fill_manual(values = my.color.neural[c(1, 3, 5, 7, 9, 11)]) +
  theme(legend.position = "none", axis.text = element_text(colour = "black")) +
    xlab("")



```




## Supp 4E: Feature plot of pbx3 and meis1 in all cells
```{r supp 4d, cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE}

# set assay
DefaultAssay(pax.all) <- "RNA"
Idents(pax.all) <- "AK_Names"

p1 <- FeaturePlot(pax.all, features = c("pbx3"),  order = TRUE, cols = c("gold", "green", "purple"))+ NoAxes()+NoLegend()
p2 <- FeaturePlot(pax.all, features = c("meis1"),  order = TRUE, cols = c("gold", "green", "purple"))+ NoAxes()+NoLegend()

plot_grid(p2, p1, nrow = 2)
```

## Supp 3F: Aztekin et al neural cell distinction
```{r aztekin, cache=TRUE, warning=FALSE, message=FALSE, echo=FALSE}

Idents(pax.neural) <- pax.neural$Laevis

neuralgenes <- c("mbp", "lepr", "lep", "foxa1", "foxa2", "tph1", "th", "gata3", "gata2", "tubb3", "mmp17", "mmp15", "tlx3", "tlx1", "pou4f1", "vsx2", "chat", "isl1", "mnx1", "lhx1", "pax2", "evx2", "evx1", "neurod4", "neurod1", "ascl1", "olig2", "dbx1", "irx3", "nkx6-1", "pax6", "nkx6-2", "arx")

DefaultAssay(pax.neural) <- "RNA"
dot <- (DotPlot(pax.neural, features=neuralgenes))$data
ggplot(dot, aes(x=features.plot, y=id)) +
  geom_point(aes(size=pct.exp, color=avg.exp.scaled)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text = element_text(colour = "black", size = 30), axis.text.x = element_text(color="black"), axis.text.y = element_text(color="black"))+
  scale_color_gradient2(low = "gold", mid = "green", high = "purple", limits = c(0, 2.5)) +
  xlab("") +
  ylab("") +
  Seurat:::RotatedAxis() +
  Seurat:::NoLegend()

```




